---
title: "Subjective-discount calculations"
output: html_document
---

This file contains only the calculations needed for the subjective-discount
exercise and the two figures used in the paper.

The treated sample contains EU countries other than Ireland. A student entering
in 2021 completes a standard three-year programme in 2024 and becomes liable to
repay from April 2025. The loan thresholds are therefore the official 2025-26
Plan 2 overseas thresholds, applicable from April 2025 to March 2026. Observed
2024 wages are converted from euros to pounds using the official 2025-26 SLC
rate of 0.848248 and grown once at the assumed 2% rate to obtain earnings in the
first repayment year.

Official source
https://www.gov.uk/government/publications/overseas-earnings-thresholds-for-plan-2-student-loans/overseas-earnings-thresholds-for-plan-2-student-loans-2024-25

```{r setup, message=FALSE, warning=FALSE}
library(readxl)
library(lmtest)
library(sandwich)

loan_amount <- 9250
post_brexit_price <- 22200
repayment_rate <- 0.09
wage_growth <- 0.02
years_to_first_repayment <- 1
rpi <- 0.043
interest_spread <- 0.03
eur_gbp <- 0.848248
```

# Data

```{r thresholds}
# https://www.gov.uk/government/publications/overseas-earnings-thresholds-for-plan-2-student-loans/overseas-earnings-thresholds-for-plan-2-student-loans-2024-25

thresholds <- data.frame(
  Country = c(
    "Austria", "Belgium", "Bulgaria", "Croatia", "Cyprus", "Czechia",
    "Denmark", "Estonia", "Finland", "France", "Germany", "Greece",
    "Hungary", "Italy", "Latvia", "Lithuania", "Luxembourg", "Malta",
    "Netherlands", "Poland", "Portugal", "Romania", "Slovakia",
    "Slovenia", "Spain", "Sweden"
  ),
  iso2c = c(
    "AT", "BE", "BG", "HR", "CY", "CZ", "DK", "EE", "FI", "FR",
    "DE", "GR", "HU", "IT", "LV", "LT", "LU", "MT", "NL", "PL",
    "PT", "RO", "SK", "SI", "ES", "SE"
  ),
  lower = c(
    28470, 28470, 11390, 17085, 22780, 17085, 28470, 22780, 28470,
    22780, 28470, 17085, 17085, 22780, 17085, 17085, 28470, 22780,
    28470, 17085, 17085, 11390, 17085, 22780, 22780, 28470
  ),
  upper = c(
    51245, 51245, 20500, 30750, 41000, 30750, 51245, 41000, 51245,
    41000, 51245, 30750, 30750, 41000, 30750, 30750, 51245, 41000,
    51245, 30750, 30750, 20500, 30750, 41000, 41000, 51245
  )
)
```

```{r inputs}
wages <- read_excel("wages.xlsx", sheet = "main_df")
betas <- read_excel("betas_saturated_model.xlsx", sheet = "main")

# Remove additions such as "(European Union)" from country names.
wages$Country <- trimws(sub("\\s*\\([^()]*\\)\\s*$", "", wages$Country))
betas$Country <- trimws(sub("\\s*\\([^()]*\\)\\s*$", "", betas$Country))

# Use one country name everywhere.
wages$Country[wages$Country == "Czech Republic"] <- "Czechia"
betas$Country[betas$Country == "Czech Republic"] <- "Czechia"

wages <- wages[, c("Country", "Wage_EURO")]
betas <- betas[, c(
  "Country",
  "Estimate_TWFE",
  "Estimate_controls",
  "Estimate_Pooled"
)]

dat <- merge(thresholds, wages, by = "Country")
dat <- merge(dat, betas, by = "Country")

dat$wage_gbp_2024 <- dat$Wage_EURO * eur_gbp

# Repayments begin in April 2025, one year after the observed wage reference.
dat$wage_gbp <- dat$wage_gbp_2024 *
  (1 + wage_growth)^years_to_first_repayment

# Clear names for the three coefficient series.
dat$beta_TWFE_no_controls <- as.numeric(dat$Estimate_TWFE)
dat$beta_TWFE_controls <- as.numeric(dat$Estimate_controls)
dat$beta_pooled <- as.numeric(dat$Estimate_Pooled)

dat <- dat[order(dat$iso2c), ]
row.names(dat) <- NULL

dat[, c(
  "iso2c",
  "Country",
  "beta_TWFE_no_controls",
  "beta_TWFE_controls",
  "beta_pooled",
  "Wage_EURO",
  "wage_gbp_2024",
  "wage_gbp",
  "lower",
  "upper"
)]
```

# GDP per capita <<<<

Only the World Bank GDP-per-capita series is needed. If
`wdi_snapshot.csv` already exists, the file is reused. Otherwise, the series is
downloaded once and saved.

```{r gdp}
if (file.exists("wdi_snapshot.csv")) {
  gdp <- read.csv("wdi_snapshot.csv")
} else {
  gdp <- WDI::WDI(
    country = dat$iso2c,
    indicator = c(GDP_PPP_pc = "NY.GDP.PCAP.PP.KD"), # check this if the same
    start = 2018,
    end = 2022,
    extra = FALSE
  )
  write.csv(gdp, "wdi_snapshot.csv", row.names = FALSE)
}

gdp <- gdp[!is.na(gdp$GDP_PPP_pc), ]
gdp <- gdp[order(gdp$iso2c, -gdp$year), ]
gdp <- gdp[!duplicated(gdp$iso2c), c("iso2c", "GDP_PPP_pc")]

dat <- merge(dat, gdp, by = "iso2c")
dat$log_gdp <- log(dat$GDP_PPP_pc)
dat <- dat[order(dat$iso2c), ]
row.names(dat) <- NULL
```

# Which estimate is used

Every subjective-discount elasticity below uses `beta_TWFE_controls`. This is
the country-specific TWFE estimate from the specification with controls.

`beta_TWFE_no_controls` is the country-specific TWFE estimate without
controls. `beta_pooled` is the pooled estimate. These two series are shown only
in the final raw-coefficient figure. They are not used to construct the main
elasticities.

```{r estimate-choice}
beta_used_in_elasticities <- dat$beta_TWFE_controls
```


#intergeneartional wealth 

```{r}
library(haven)
library(countrycode)

gdim_file <- "gdim_income.dta"

if (!file.exists(gdim_file)) {
  download.file(
    "https://datacatalogfiles.worldbank.org/ddh-published/0066878/DR0095414/IGE_Munoz_VanderWeide_June2025.dta",
    gdim_file,
    mode = "wb"
  )
}

mobility <- read_dta(gdim_file)

mobility$iso2c <- countrycode(
  mobility$code,
  origin = "iso3c",
  destination = "iso2c"
)

mobility <- mobility[, c("iso2c", "IGE")]

dat <- merge(
  dat,
  mobility,
  by = "iso2c",
  all.x = TRUE
)

dat <- dat[order(dat$iso2c), ]
row.names(dat) <- NULL

# Higher IGE means stronger income persistence and lower mobility
dat[, c("iso2c", "Country", "IGE")]

stopifnot(!any(is.na(dat$IGE)))
```




# Subjective value of the loan

This is the only calculation function in the file. It simulates repayments for
one country and returns their subjective present value.

```{r loan-value}
loan_pv <- function(wage, lower, upper, rho) {

  balance <- loan_amount
  pv <- 0

  for (t in 1:30) {

    earnings <- wage * (1 + wage_growth)^(t - 1)

    if (earnings <= lower) loading <- 0
    if (earnings > lower & earnings < upper) {
      loading <- (earnings - lower) / (upper - lower)
    }
    if (earnings >= upper) loading <- 1

    balance <- balance * (1 + rpi + interest_spread * loading)
    payment <- min(repayment_rate * max(0, earnings - lower), balance)

    pv <- pv + payment / (1 + rho)^t
    balance <- balance - payment

    if (balance <= 0) break
  }

  pv
}
```

# Find the two discount-rate thresholds

The search runs from 3% to 35%. At each rate, the code calculates the
controlled-TWFE elasticities, regresses them on log GDP per capita, and stores
the income-gradient coefficient and its HC1 p-value.

```{r thresholds-search}
rho_grid <- seq(0.03, 0.35, by = 0.001)

elasticity_grid <- matrix(
  NA_real_,
  nrow = nrow(dat),
  ncol = length(rho_grid)
)

income_gradient <- numeric(length(rho_grid))
income_p_value <- numeric(length(rho_grid))

for (j in seq_along(rho_grid)) {

  pv <- mapply(
    
    loan_pv,
    
    dat$wage_gbp,
    
    dat$lower,
    
    dat$upper,
    
    MoreArgs = list(rho = rho_grid[j])
    
  )

  experienced_price_change <- log(post_brexit_price) - log(pv)

  elasticity_grid[, j] <- beta_used_in_elasticities / experienced_price_change

  model <- lm(elasticity_grid[, j] ~ dat$log_gdp)

  income_gradient[j] <- coef(model)[2]
  
  income_p_value[j] <- coeftest(
    model,
    vcov. = vcovHC(model, type = "HC1")
  )[2, 4]
}

rho_insignificant <- rho_grid[which(income_p_value >= 0.05)[1]]
rho_neutral <- rho_grid[which.min(abs(income_gradient))]

cat("First grid rate with an insignificant income gradient ",
    round(100 * rho_insignificant, 1), "%\n", sep = "")
cat("Income-neutral discount rate ",
    round(100 * rho_neutral, 1), "%\n", sep = "")
```

# EXAMPLE IN TEXT

```{r}

austria <- dat[dat$Country == "Austria", ]
bulgaria <- dat[dat$Country == "Bulgaria", ]

austria_pv_3 <- loan_pv(
  austria$wage_gbp,
  austria$lower,
  austria$upper,
  0.03
)

bulgaria_pv_3 <- loan_pv(
  bulgaria$wage_gbp,
  bulgaria$lower,
  bulgaria$upper,
  0.03
)

austria_log_price_change_3 <-
  log(post_brexit_price) - log(austria_pv_3)

bulgaria_log_price_change_3 <-
  log(post_brexit_price) - log(bulgaria_pv_3)

c(
  austria_wage_gbp_2025 = austria$wage_gbp,
  austria_pv_3 = austria_pv_3,
  austria_log_price_change_3 = austria_log_price_change_3,
  bulgaria_wage_gbp_2025 = bulgaria$wage_gbp,
  bulgaria_pv_3 = bulgaria_pv_3,
  bulgaria_log_price_change_3 = bulgaria_log_price_change_3
)
```



# Figure 5 Panel C

The orange point uses the controlled-TWFE elasticity when the income gradient
first becomes insignificant. The blue point uses the controlled-TWFE
elasticity at income neutrality. The green point uses the same controlled-TWFE
coefficient at a 35% subjective discount rate. The shaded interval shows the
literature range from -0.80 to -0.10.

```{r country-elasticities, fig.width=8, fig.height=8.5}
eta_insignificant <- elasticity_grid[
  , which.min(abs(rho_grid - rho_insignificant))
]

eta_neutral <- elasticity_grid[
  , which.min(abs(rho_grid - rho_neutral))
]

eta_35 <- elasticity_grid[, ncol(elasticity_grid)]

elasticities <- cbind(
  eta_insignificant,
  eta_neutral,
  eta_35
)

literature_bounds <- c(-0.80, -0.10)
country_order <- order(rowMeans(elasticities))
y <- seq_len(nrow(dat))

par(mar = c(8, 4, 5, 2))

plot(
  range(c(elasticities, literature_bounds)),
  range(y),
  type = "n",
  yaxt = "n",
  xlab = "Estimated price elasticity",
  ylab = ""
)

axis(
  2,
  at = y,
  labels = dat$iso2c[country_order],
  las = 1,
  tick = FALSE
)

rect(
  literature_bounds[1],
  par("usr")[3],
  literature_bounds[2],
  par("usr")[4],
  col = adjustcolor("grey70", alpha.f = 0.3),
  border = NA
)

abline(v = 0, col = "grey60")

segments(
  apply(elasticities[country_order, ], 1, min),
  y,
  apply(elasticities[country_order, ], 1, max),
  y,
  col = "#9FB4C7",
  lwd = 4
)

points(
  eta_insignificant[country_order], y,
  pch = 16, col = "#D55E00"
)

points(
  eta_neutral[country_order], y,
  pch = 16, col = "#1F4E79"
)

points(
  eta_35[country_order], y,
  pch = 16, col = "#009E73"
)

title(
  "Country elasticities under alternative discount rates",
  line = 2
)

mtext(
  paste0(
    "Insignificance at ", round(100 * rho_insignificant, 1),
    "%, income neutrality at ", round(100 * rho_neutral, 1),
    "%, and upper benchmark at 35%"
  ),
  side = 3,
  line = 0.5
)

legend(
  "bottom",
  inset = c(0, -0.32),
  xpd = NA,
  ncol = 2,
  legend = c(
    "Insignificance threshold",
    "Income neutrality",
    "35% discount rate",
    "Literature bounds"
  ),
  col = c(
    "#D55E00",
    "#1F4E79",
    "#009E73",
    adjustcolor("grey70", alpha.f = 0.6)
  ),
  pch = c(16, 16, 16, 15),
  pt.cex = c(1, 1, 1, 2),
  cex = 0.8,
  bty = "n"
)
```

# Figure 5 Panel B

This figure uses the raw country coefficients from the beta workbook. No
subjective present-value correction is applied. GDP per capita is also shown in
levels instead of logs. The three panels therefore display the baseline
country correlations before the elasticity calibration.

```{r raw-correlations, fig.width=9, fig.height=4.5}
raw_estimates <- cbind(
  dat$beta_TWFE_controls,
  dat$beta_TWFE_no_controls,
  dat$beta_pooled
)

panel_names <- c(
  "TWFE with controls",
  "TWFE without controls",
  "Pooled"
)

panel_colours <- c("#1F4E79", "#D55E00", "#009E73")
log_gdp <- dat$log_gdp

par(
  mfrow = c(1, 3),
  oma = c(4, 4, 3, 1),
  mar = c(2, 2, 3, 1)
)

for (j in 1:3) {

  plot(
    log_gdp,
    raw_estimates[, j],
    pch = 16,
    col = panel_colours[j],
    xlab = "",
    ylab = "",
    main = panel_names[j]
  )

  text(
    log_gdp,
    raw_estimates[, j],
    labels = dat$iso2c,
    pos = 3,
    cex = 0.65,
    col = panel_colours[j]
  )

  abline(
    lm(raw_estimates[, j] ~ log_gdp),
    col = "red",
    lwd = 2
  )

  mtext(
    paste0("Correlation = ",
           round(cor(log_gdp, raw_estimates[, j]), 2)),
    side = 3,
    line = 0.2,
    cex = 0.75
  )
}

mtext(
  "GDP per capita, PPP, constant, thousands",
  side = 1,
  outer = TRUE,
  line = 2.2
)

mtext(
  "Raw country coefficient",
  side = 2,
  outer = TRUE,
  line = 2.2
)

mtext(
  "Raw treatment coefficients and log of GDP per capita",
  side = 3,
  outer = TRUE,
  line = 1.3,
  cex = 1.1
)

par(mfrow = c(1, 1))
```

# Figure 5 Panel A

```{r}
three_elasticities <- cbind(
  elasticity_grid[, 1],
  eta_neutral,
  eta_35
)

three_titles <- c(
  "rho = 3.0%",
  paste0("gamma = 0\nrho = ", round(100 * rho_neutral, 1), "%"),
  "rho = 35.0%"
)

par(
  mfrow = c(1, 3),
  oma = c(4, 4, 4, 1),
  mar = c(2, 2, 3, 1)
)

for (j in 1:3) {

  plot(
    dat$log_gdp,
    three_elasticities[, j],
    pch = 16,
    col = "#1F4E79",
    xlab = "",
    ylab = "",
    main = three_titles[j]
  )

  text(
    dat$log_gdp,
    three_elasticities[, j],
    labels = dat$iso2c,
    pos = 3,
    cex = 0.65,
    col = "#1F4E79"
  )

  abline(
    lm(three_elasticities[, j] ~ dat$log_gdp),
    col = "red",
    lwd = 2
  )
}

mtext(
  "log GDP per capita, PPP, constant",
  side = 1,
  outer = TRUE,
  line = 2.2
)

mtext(
  expression(eta[j]),
  side = 2,
  outer = TRUE,
  line = 2.2
)

mtext(
  "Country elasticities at 3%, income neutrality, and 35%",
  side = 3,
  outer = TRUE,
  line = 1.5,
  cex = 1.15
)

par(mfrow = c(1, 1))
```


```{r}
# All elasticities use the controlled-TWFE coefficients

rho_alt <- seq(0.03, 0.40, by = 0.005)

eta_alt <- matrix(
  NA_real_,
  nrow = nrow(dat),
  ncol = length(rho_alt)
)

for (j in seq_along(rho_alt)) {

  pv <- mapply(
    loan_pv,
    dat$wage_gbp,
    dat$lower,
    dat$upper,
    MoreArgs = list(rho = rho_alt[j])
  )

  eta_alt[, j] <- beta_used_in_elasticities /
    (log(post_brexit_price) - log(pv))
}

wdi <- read.csv("wdi_snapshot.csv")

wdi$log_cons <- log(wdi$HH_cons_pc)
wdi$log_gdp <- log(wdi$GDP_PPP_pc)
wdi$log_gni <- log(wdi$GNI_PPP_pc)
wdi$log_pop <- log(wdi$Population)

covariate_names <- c(
  "Educ_spend_gdp",
  "GDP_growth",
  "Gini",
  "Inflation",
  "log_cons",
  "log_gdp",
  "log_gni",
  "log_pop",
  "Tertiary_attain",
  "Tertiary_enrol",
  "Unemp",
  "Youth_unemp"
)

panel_names <- c(
  "Education spending",
  "GDP growth",
  "Gini",
  "Inflation",
  "Consumption per capita",
  "GDP per capita",
  "GNI per capita",
  "Population",
  "Tertiary attainment",
  "Tertiary enrolment",
  "Unemployment",
  "Youth unemployment"
)

covariates <- matrix(
  NA_real_,
  nrow = nrow(dat),
  ncol = length(covariate_names)
)

for (v in seq_along(covariate_names)) {

  for (i in seq_len(nrow(dat))) {

    country_data <- wdi[
      wdi$iso2c == dat$iso2c[i] &
        !is.na(wdi[[covariate_names[v]]]),
    ]

    if (nrow(country_data) > 0) {

      latest <- which.max(country_data$year)

      covariates[i, v] <-
        country_data[[covariate_names[v]]][latest]
    }
  }
}

# Add intergenerational income persistence
# Higher IGE means lower intergenerational mobility

covariates <- cbind(
  covariates,
  IGE = as.numeric(dat$IGE)
)

covariate_names <- c(
  covariate_names,
  "IGE"
)

panel_names <- c(
  panel_names,
  "Intergenerational\nincome persistence"
)

# Correlations and HC1 confidence intervals

correlations <- matrix(
  NA_real_,
  nrow = length(rho_alt),
  ncol = length(covariate_names)
)

ci_lower <- correlations
ci_upper <- correlations

for (v in seq_along(covariate_names)) {

  for (j in seq_along(rho_alt)) {

    ok <- is.finite(covariates[, v]) &
      is.finite(eta_alt[, j])

    x <- covariates[ok, v]
    y <- eta_alt[ok, j]

    model <- lm(scale(y) ~ scale(x))

    robust_result <- coeftest(
      model,
      vcov. = vcovHC(model, type = "HC1")
    )

    correlations[j, v] <- robust_result[2, 1]

    ci_lower[j, v] <- robust_result[2, 1] -
      1.96 * robust_result[2, 2]

    ci_upper[j, v] <- robust_result[2, 1] +
      1.96 * robust_result[2, 2]
  }
}

# Plot

par(
  mfrow = c(4, 4),
  oma = c(4, 4, 5, 1),
  mar = c(2, 2.5, 2.5, 1)
)

for (v in seq_along(covariate_names)) {

  plot(
    100 * rho_alt,
    correlations[, v],
    type = "n",
    xaxt = "n",
    xlab = "",
    ylab = "",
    ylim = range(
      ci_lower[, v],
      ci_upper[, v],
      na.rm = TRUE
    ),
    main = panel_names[v]
  )

  polygon(
    c(100 * rho_alt, rev(100 * rho_alt)),
    c(ci_lower[, v], rev(ci_upper[, v])),
    col = adjustcolor("#1F4E79", alpha.f = 0.15),
    border = NA
  )

  grid(col = "grey90")
  abline(h = 0, col = "grey60")

  lines(
    100 * rho_alt,
    correlations[, v],
    col = "#1F4E79",
    lwd = 2
  )

  axis(
    1,
    at = c(3, 10, 20, 30, 40),
    labels = paste0(c(3, 10, 20, 30, 40), "%")
  )

  sign_change <- which(
    correlations[-nrow(correlations), v] *
      correlations[-1, v] < 0
  )[1]

  if (!is.na(sign_change)) {

    abline(
      v = 100 * rho_alt[sign_change],
      col = "red",
      lwd = 2
    )
  }
}

mtext(
  "Subjective discount rate",
  side = 1,
  outer = TRUE,
  line = 2.2
)

mtext(
  expression("Correlation with " * eta[j]),
  side = 2,
  outer = TRUE,
  line = 2.2
)

mtext(
  "Which correlations survive the discounting assumption",
  side = 3,
  outer = TRUE,
  line = 2.5,
  cex = 1.2
)

mtext(
  "Shaded bands show 95% HC1 confidence intervals; red line marks a sign change",
  side = 3,
  outer = TRUE,
  line = 1.2,
  cex = 0.85
)

par(mfrow = c(1, 1))
```
