---
title: "Subjective-discount calculations, fixed-RPI 2025 calibration"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

This file contains only the calculations needed for the
subjective-discount exercise and the two figures used in the paper.

The treated sample contains EU countries other than Ireland. A student
entering in 2021 completes a standard three-year programme in 2024 and
becomes liable to repay from April 2025. The loan thresholds are
therefore the official 2025-26 Plan 2 overseas thresholds, applicable
from April 2025 to March 2026. Observed 2024 wages are converted from
euros to pounds using the official 2025-26 SLC rate of 0.848248 and
grown once at the assumed 2% rate to obtain earnings in the first
repayment year.

Official source
<https://www.gov.uk/government/publications/overseas-earnings-thresholds-for-plan-2-student-loans/overseas-earnings-thresholds-for-plan-2-student-loans-2024-25>

Official interest sources
<https://www.gov.uk/guidance/how-interest-is-calculated-plan-2>
<https://www.gov.uk/government/news/student-loans-interest-rates-and-repayment-threshold-announcement--5>
<https://www.gov.uk/government/news/student-loans-interest-rates-and-repayment-threshold-announcement--6>

```{r setup, message=FALSE, warning=FALSE}
library(readxl)
library(lmtest)
library(sandwich)

loan_amount <- 9250
post_brexit_price <- 22200
repayment_rate <- 0.09
wage_growth <- 0.02
years_to_first_repayment <- 1

# The first repayment financial year runs from April 2025 to March 2026.
# It contains five months under the 2024-25 RPI rate and seven months under
# the 2025-26 RPI rate.
first_fy_rpi_schedule <- c(0.043, 0.032)
first_fy_month_weights <- c(5, 7)
interest_spread <- 0.03
eur_gbp <- 0.848248

stopifnot(sum(first_fy_month_weights) == 12)

# Derive one fixed RPI base, then retain the standard 0% to 3% income loading.
fixed_rpi <- weighted.mean(
  first_fy_rpi_schedule,
  w = first_fy_month_weights
)

period_interest_rate <- function(loading) {
  fixed_rpi + interest_spread * loading
}
```

# Data

```{r thresholds}
thresholds <- data.frame(
  Country = c(
    "Austria", "Belgium", "Bulgaria", "Croatia", "Cyprus", "Czechia",
    "Denmark", "Estonia", "Finland", "France", "Germany", "Greece",
    "Hungary", "Italy", "Latvia", "Lithuania", "Luxembourg", "Malta",
    "Netherlands", "Poland", "Portugal", "Romania", "Slovakia",
    "Slovenia", "Spain", "Sweden"
  ),
  iso2c = c(
    "AT", "BE", "BG", "HR", "CY", "CZ", "DK", "EE", "FI", "FR",
    "DE", "GR", "HU", "IT", "LV", "LT", "LU", "MT", "NL", "PL",
    "PT", "RO", "SK", "SI", "ES", "SE"
  ),
  lower = c(
    28470, 28470, 11390, 17085, 22780, 17085, 28470, 22780, 28470,
    22780, 28470, 17085, 17085, 22780, 17085, 17085, 28470, 22780,
    28470, 17085, 17085, 11390, 17085, 22780, 22780, 28470
  ),
  upper = c(
    51245, 51245, 20500, 30750, 41000, 30750, 51245, 41000, 51245,
    41000, 51245, 30750, 30750, 41000, 30750, 30750, 51245, 41000,
    51245, 30750, 30750, 20500, 30750, 41000, 41000, 51245
  )
)
```

```{r inputs}
wages <- read_excel("wages.xlsx", sheet = "main_df")
betas <- read_excel("betas_saturated_model.xlsx", sheet = "main")

# Remove additions such as "(European Union)" from country names.
wages$Country <- trimws(sub("\\s*\\([^()]*\\)\\s*$", "", wages$Country))
betas$Country <- trimws(sub("\\s*\\([^()]*\\)\\s*$", "", betas$Country))

# Use one country name everywhere.
wages$Country[wages$Country == "Czech Republic"] <- "Czechia"
betas$Country[betas$Country == "Czech Republic"] <- "Czechia"

wages <- wages[, c("Country", "Wage_EURO")]
betas <- betas[, c(
  "Country",
  "Estimate_TWFE",
  "Estimate_controls",
  "Estimate_Pooled"
)]

dat <- merge(thresholds, wages, by = "Country")
dat <- merge(dat, betas, by = "Country")

dat$wage_gbp_2024 <- dat$Wage_EURO * eur_gbp

# Repayments begin in April 2025, one year after the observed wage reference.
dat$wage_gbp <- dat$wage_gbp_2024 *
  (1 + wage_growth)^years_to_first_repayment

# Clear names for the three coefficient series.
dat$beta_TWFE_no_controls <- as.numeric(dat$Estimate_TWFE)
dat$beta_TWFE_controls <- as.numeric(dat$Estimate_controls)
dat$beta_pooled <- as.numeric(dat$Estimate_Pooled)

dat <- dat[order(dat$iso2c), ]
row.names(dat) <- NULL

dat[, c(
  "iso2c",
  "Country",
  "beta_TWFE_no_controls",
  "beta_TWFE_controls",
  "beta_pooled",
  "Wage_EURO",
  "wage_gbp_2024",
  "wage_gbp",
  "lower",
  "upper"
)]
```

# GDP per capita \<\<\<\<

Only the World Bank GDP-per-capita series is needed. If
`wdi_snapshot.csv` already exists, the file is reused. Otherwise, the
series is downloaded once and saved.

```{r gdp}
if (file.exists("wdi_snapshot.csv")) {
  gdp <- read.csv("wdi_snapshot.csv")
} else {
  gdp <- WDI::WDI(
    country = dat$iso2c,
    indicator = c(GDP_PPP_pc = "NY.GDP.PCAP.PP.KD"), # check this if the same
    start = 2018,
    end = 2022,
    extra = FALSE
  )
  write.csv(gdp, "wdi_snapshot.csv", row.names = FALSE)
}

gdp <- gdp[!is.na(gdp$GDP_PPP_pc), ]
gdp <- gdp[order(gdp$iso2c, -gdp$year), ]
gdp <- gdp[!duplicated(gdp$iso2c), c("iso2c", "GDP_PPP_pc")]

dat <- merge(dat, gdp, by = "iso2c")
dat$log_gdp <- log(dat$GDP_PPP_pc)
dat <- dat[order(dat$iso2c), ]
row.names(dat) <- NULL
```

# Which estimate is used

Every subjective-discount elasticity below uses `beta_TWFE_controls`.
This is the country-specific TWFE estimate from the specification with
controls.

`beta_TWFE_no_controls` is the country-specific TWFE estimate without
controls. `beta_pooled` is the pooled estimate. These two series are
shown only in the final raw-coefficient figure. They are not used to
construct the main elasticities.

```{r estimate-choice}
beta_used_in_elasticities <- dat$beta_TWFE_controls
```

# intergenerational wealth

```{r}
library(haven)
library(countrycode)

gdim_file <- "gdim_income.dta"

if (!file.exists(gdim_file)) {
  download.file(
    "https://datacatalogfiles.worldbank.org/ddh-published/0066878/DR0095414/IGE_Munoz_VanderWeide_June2025.dta",
    gdim_file,
    mode = "wb"
  )
}

mobility <- read_dta(gdim_file)

mobility$iso2c <- countrycode(
  mobility$code,
  origin = "iso3c",
  destination = "iso2c"
)

mobility <- mobility[, c("iso2c", "IGE")]

dat <- merge(
  dat,
  mobility,
  by = "iso2c",
  all.x = TRUE
)

dat <- dat[order(dat$iso2c), ]
row.names(dat) <- NULL

# Higher IGE means stronger income persistence and lower mobility
dat[, c("iso2c", "Country", "IGE")]

stopifnot(!any(is.na(dat$IGE)))
```

# Subjective value of the loan

This is the only calculation function in the file. It simulates
repayments for one country and returns their subjective present value.

For earnings $Y_{jt}$, lower threshold $\tau_j^L$, and upper threshold
$\tau_j^U$, the income loading is

$$
\ell_{jt}=
\begin{cases}
0, & Y_{jt}\leq\tau_j^L,\\
\dfrac{Y_{jt}-\tau_j^L}{\tau_j^U-\tau_j^L},
& \tau_j^L<Y_{jt}<\tau_j^U,\\
1, & Y_{jt}\geq\tau_j^U.
\end{cases}
$$

The first repayment financial year crosses two statutory RPI periods.
Their month-weighted average defines the fixed RPI base

$$
\overline{RPI}^{2025/26}
=\frac{5}{12}(0.043)+\frac{7}{12}(0.032)
=0.0365833.
$$

The standard income-dependent spread is then added without modification

$$
i_{jt}^{2025/26}
=\overline{RPI}^{2025/26}+0.03\ell_{jt}
=0.0365833+0.03\ell_{jt}.
$$

The previous code used $0.043+0.03\ell_{jt}$. The correction leaves the
income loading unchanged and lowers the common RPI component by 0.6417
percentage points. To preserve a like-for-like sensitivity, the
first-year equivalent is held fixed in all 30 annual iterations. This is
a stationary calibration, not a forecast of future RPI or thresholds.
The opening loan amount is interpreted as the balance at the statutory
repayment due date, so interest accrued during study is outside this
calculation.

The remaining annual-frequency equations exactly mirror the R function

$$
Y_{jt}=w_j(1+g)^{t-1},
$$

$$
\widetilde B_{jt}=B_{j,t-1}(1+i_{jt}),
$$

$$
P_{jt}=\min\left\{0.09\max(0,Y_{jt}-\tau_j^L),
\widetilde B_{jt}\right\},
\qquad
B_{jt}=\widetilde B_{jt}-P_{jt},
$$

$$
PV_j(\rho)=\sum_{t=1}^{30}\frac{P_{jt}}{(1+\rho)^t},
\qquad B_{j0}=9250,
$$

$$
k_j(\rho)=\log(22200)-\log\!\left(PV_j(\rho)\right),
\qquad
\eta_j(\rho)=\frac{\widehat\beta_j}{k_j(\rho)}.
$$

The code holds the country thresholds fixed and grows earnings by
$g=0.02$ each year.

```{r loan-value}
loan_pv <- function(wage, lower, upper, rho) {

  balance <- loan_amount
  pv <- 0

  for (t in 1:30) {

    earnings <- wage * (1 + wage_growth)^(t - 1)

    if (earnings <= lower) loading <- 0
    if (earnings > lower & earnings < upper) {
      loading <- (earnings - lower) / (upper - lower)
    }
    if (earnings >= upper) loading <- 1

    applied_interest_rate <- period_interest_rate(loading)
    balance <- balance * (1 + applied_interest_rate)
    payment <- min(repayment_rate * max(0, earnings - lower), balance)

    pv <- pv + payment / (1 + rho)^t
    balance <- balance - payment

    if (balance <= 0) break
  }

  pv
}
```

# Find the two discount-rate thresholds

The search runs from 3% to 35%. At each rate, the code calculates the
controlled-TWFE elasticities, regresses them on log GDP per capita, and
stores the income-gradient coefficient and its HC1 p-value.

```{r thresholds-search}
rho_grid <- seq(0.03, 0.35, by = 0.001)

elasticity_grid <- matrix(
  
  NA_real_,
  
  nrow = nrow(dat),
  
  ncol = length(rho_grid)
)

income_gradient <- numeric(length(rho_grid))

income_p_value <- numeric(length(rho_grid))

for (j in seq_along(rho_grid)) {
  
  pv <- mapply( loan_pv,dat$wage_gbp,dat$lower,dat$upper,MoreArgs = list(rho = rho_grid[j]))

  experienced_price_change <- log(post_brexit_price) - log(pv)

  elasticity_grid[, j] <- beta_used_in_elasticities / experienced_price_change

  model <- lm(elasticity_grid[, j] ~ dat$log_gdp)

  income_gradient[j] <- coef(model)[2]
  
  income_p_value[j] <- coeftest( model, vcov. = vcovHC(model, type = "HC1") )[2, 4]
}

rho_insignificant <- rho_grid[which(income_p_value >= 0.05)[1]]

rho_neutral <- rho_grid[which.min(abs(income_gradient))]

cat("First grid rate with an insignificant income gradient ",
    
    round(100 * rho_insignificant, 1), "%\n", sep = "")

cat("Income-neutral discount rate ",
    
    round(100 * rho_neutral, 1), "%\n", sep = "")
```

# EXAMPLE IN TEXT

```{r}

austria <- dat[dat$Country == "Austria", ]
bulgaria <- dat[dat$Country == "Bulgaria", ]

austria_pv_3 <- loan_pv(
  austria$wage_gbp,
  austria$lower,
  austria$upper,
  0.03
)

bulgaria_pv_3 <- loan_pv(
  bulgaria$wage_gbp,
  bulgaria$lower,
  bulgaria$upper,
  0.03
)

bulgaria_pv_149 <- loan_pv(
  bulgaria$wage_gbp,
  bulgaria$lower,
  bulgaria$upper,
  0.149
)

austria_log_price_change_3 <-
  log(post_brexit_price) - log(austria_pv_3)

bulgaria_log_price_change_3 <-
  log(post_brexit_price) - log(bulgaria_pv_3)

bulgaria_pv_3 
bulgaria_pv_149

luxembourg_3<-loan_pv(
  wages[wages$Country=="Luxembourg",]$Wage_EURO*0.848248*(1+0.02),
  thresholds[thresholds$Country=="Luxembourg",]$lower,
  thresholds[thresholds$Country=="Luxembourg",]$upper,
  0.03
)
luxembourg_3

portugal_3pct<- loan_pv(wages[wages$Country=="Portugal",]$Wage_EURO*0.848248*(1+0.02), thresholds[thresholds$Country=="Portugal",]$lower,thresholds[thresholds$Country=="Portugal",]$upper, 0.03 )

portugal_3pct

betas[betas$Country=="Portugal",]$Estimate_controls/(log(22200)-log(portugal_3pct))


```

# Figure 5 Panel C

The orange point uses the controlled-TWFE elasticity when the income
gradient first becomes insignificant. The blue point uses the
controlled-TWFE elasticity at income neutrality. The green point uses
the same controlled-TWFE coefficient at a 35% subjective discount rate.
The shaded interval shows the literature range from -0.80 to -0.10.

```{r country-elasticities, fig.width=8, fig.height=7}
eta_insignificant <- elasticity_grid[ , which.min(abs(rho_grid - rho_insignificant))]

eta_neutral <- elasticity_grid[, which.min(abs(rho_grid - rho_neutral))]

eta_35 <- elasticity_grid[, ncol(elasticity_grid)]

elasticities <- cbind(
  eta_insignificant,
  eta_neutral,
  eta_35
)

literature_bounds <- c(-0.80, -0.10)

country_order <- order(rowMeans(elasticities))

y <- seq_len(nrow(dat))

par(mar = c(8, 4, 5, 2))

plot(
  range(c(elasticities, literature_bounds)),
  range(y),
  type = "n",
  yaxt = "n",
  xlab = "Estimated price elasticity",
  ylab = ""
)

axis(
  2,
  at = y,
  labels = dat$iso2c[country_order],
  las = 1,
  tick = FALSE
)

rect(
  literature_bounds[1],
  par("usr")[3],
  literature_bounds[2],
  par("usr")[4],
  col = adjustcolor("grey70", alpha.f = 0.3),
  border = NA
)

abline(v = 0, col = "grey60")

segments(
  apply(elasticities[country_order, ], 1, min),
  y,
  apply(elasticities[country_order, ], 1, max),
  y,
  col = "#9FB4C7",
  lwd = 4
)

points(
  eta_insignificant[country_order], y,
  pch = 16, col = "#D55E00"
)

points(
  eta_neutral[country_order], y,
  pch = 16, col = "#1F4E79"
)

points(
  eta_35[country_order], y,
  pch = 16, col = "#009E73"
)

title(
  "Country elasticities under alternative discount rates",
  line = 2
)

mtext(
  paste0(
    "Insignificance at ", round(100 * rho_insignificant, 1),
    "%, income neutrality at ", round(100 * rho_neutral, 1),
    "%, and upper benchmark at 35%"
  ),
  side = 3,
  line = 0.5
)

legend(
  "bottom",
  inset = c(0, -0.32),
  xpd = NA,
  ncol = 2,
  legend = c(
    "Insignificance threshold",
    "Income neutrality",
    "35% discount rate",
    "Literature bounds"
  ),
  col = c(
    "#D55E00",
    "#1F4E79",
    "#009E73",
    adjustcolor("grey70", alpha.f = 0.6)
  ),
  pch = c(16, 16, 16, 15),
  pt.cex = c(1, 1, 1, 2),
  cex = 0.8,
  bty = "n"
)
```

# Figure 5 Panel B

This figure uses the raw country coefficients from the beta workbook. No
subjective present-value correction is applied. GDP per capita is also
shown in levels instead of logs. The three panels therefore display the
baseline country correlations before the elasticity calibration.

```{r raw-correlations, fig.width=9, fig.height=4.5}
raw_estimates <- cbind(
  dat$beta_TWFE_controls,
  dat$beta_TWFE_no_controls,
  dat$beta_pooled
)

panel_names <- c(
  "TWFE with controls",
  "TWFE without controls",
  "Pooled"
)

panel_colours <- c("#1F4E79", "#D55E00", "#009E73")
log_gdp <- dat$log_gdp

par(
  mfrow = c(1, 3),
  oma = c(4, 4, 3, 1),
  mar = c(2, 2, 3, 1)
)

for (j in 1:3) {

  plot(
    log_gdp,
    raw_estimates[, j],
    pch = 16,
    col = panel_colours[j],
    xlab = "",
    ylab = "",
    main = panel_names[j]
  )

  text(
    log_gdp,
    raw_estimates[, j],
    labels = dat$iso2c,
    pos = 3,
    cex = 0.65,
    col = panel_colours[j]
  )

  abline(
    lm(raw_estimates[, j] ~ log_gdp),
    col = "red",
    lwd = 2
  )

  mtext(
    paste0("Correlation = ",
           round(cor(log_gdp, raw_estimates[, j]), 2)),
    side = 3,
    line = 0.2,
    cex = 0.75
  )
}

mtext(
  "log (GDP per capita, PPP, constant, thousands)",
  side = 1,
  outer = TRUE,
  line = 2.2
)

mtext(
  "Raw country coefficient",
  side = 2,
  outer = TRUE,
  line = 2.2
)

mtext(
  "Raw treatment coefficients and log of GDP per capita",
  side = 3,
  outer = TRUE,
  line = 1.3,
  cex = 1.1
)

par(mfrow = c(1, 1))
```

# Figure 5 Panel A

```{r}
three_elasticities <- cbind(
  elasticity_grid[, 1],
  eta_neutral,
  eta_35
)

three_titles <- c(
  "rho = 3.0%",
  paste0("gamma = 0\nrho = ", round(100 * rho_neutral, 1), "%"),
  "rho = 35.0%"
)

par(
  mfrow = c(1, 3),
  oma = c(4, 4, 4, 1),
  mar = c(2, 2, 3, 1)
)

for (j in 1:3) {

  plot(
    dat$log_gdp,
    three_elasticities[, j],
    pch = 16,
    col = "#1F4E79",
    xlab = "",
    ylab = "",
    main = three_titles[j]
  )

  text(
    dat$log_gdp,
    three_elasticities[, j],
    labels = dat$iso2c,
    pos = 3,
    cex = 0.65,
    col = "#1F4E79"
  )

  abline(
    lm(three_elasticities[, j] ~ dat$log_gdp),
    col = "red",
    lwd = 2
  )
}

mtext(
  "log GDP per capita, PPP, constant",
  side = 1,
  outer = TRUE,
  line = 2.2
)

mtext(
  expression(eta[j]),
  side = 2,
  outer = TRUE,
  line = 2.2
)

mtext(
  "Country elasticities at 3%, income neutrality, and 35%",
  side = 3,
  outer = TRUE,
  line = 1.5,
  cex = 1.15
)

par(mfrow = c(1, 1))
```

# Correlational profiles

## Data constructions and pulls

```{r}

# All elasticities use the controlled-TWFE coefficients

rho_alt <- seq(
  0.03,
  0.40,
  by = 0.005
)

eta_alt <- matrix(
  NA_real_,
  nrow = nrow(dat),
  ncol = length(rho_alt)
)

#The elasticities matrix 

for (j in seq_along(rho_alt)) {

  pv <- mapply(loan_pv,dat$wage_gbp,dat$lower,dat$upper,MoreArgs = list(rho = rho_alt[j]))

  eta_alt[, j] <- beta_used_in_elasticities /(log(post_brexit_price) -log(pv))
}

# Load or construct the complete WDI snapshot

wdi_file <- "wdi_snapshot.csv"

wdi_indicators <- c(
  Educ_spend_gdp = "SE.XPD.TOTL.GD.ZS",
  GDP_growth = "NY.GDP.MKTP.KD.ZG",
  Gini = "SI.POV.GINI",
  Inflation = "FP.CPI.TOTL.ZG",
  HH_cons_pc = "NE.CON.PRVT.PC.KD",
  GDP_PPP_pc = "NY.GDP.PCAP.PP.KD",
  GNI_PPP_pc = "NY.GNP.PCAP.PP.KD",
  Population = "SP.POP.TOTL",
  Tertiary_attain = "SE.TER.CUAT.BA.ZS",
  Tertiary_enrol = "SE.TER.ENRR",
  Unemp = "SL.UEM.TOTL.ZS",
  Youth_unemp = "SL.UEM.1524.ZS"
)

required_wdi_columns <- c(
  "iso2c",
  "year",
  names(wdi_indicators)
)

wdi_needs_pull <- !file.exists(wdi_file)

if (!wdi_needs_pull) {

  wdi <- read.csv(
    wdi_file,
    stringsAsFactors = FALSE
  )

  wdi_needs_pull <- !all(required_wdi_columns %in% names(wdi))
}

if (wdi_needs_pull) {

  wdi <- WDI::WDI(
    country = dat$iso2c,
    indicator = wdi_indicators,
    start = 2018,
    end = 2022,
    extra = FALSE
  )

  write.csv(
    wdi,
    wdi_file,
    row.names = FALSE
  )
}

missing_wdi_columns <- setdiff(
  required_wdi_columns,
  names(wdi)
)

if (length(missing_wdi_columns) > 0) {
  stop(
    paste(
      "Missing WDI columns",
      paste(
        missing_wdi_columns,
        collapse = ", "
      )
    )
  )
}

# Construct logged covariates

wdi$log_cons <- log(wdi$HH_cons_pc)
wdi$log_gdp <- log(wdi$GDP_PPP_pc)
wdi$log_gni <- log(wdi$GNI_PPP_pc)
wdi$log_pop <- log(wdi$Population)


# Covariate definitions

covariate_names <- c(
  "Educ_spend_gdp",
  "GDP_growth",
  "Gini",
  "Inflation",
  "log_cons",
  "log_gdp",
  "log_gni",
  "log_pop",
  "Tertiary_attain",
  "Tertiary_enrol",
  "Unemp",
  "Youth_unemp"
)

panel_names <- c(
  "Education spending",
  "GDP growth",
  "Gini",
  "Inflation",
  "Consumption per capita",
  "GDP per capita",
  "GNI per capita",
  "Population",
  "Tertiary attainment",
  "Tertiary enrolment",
  "Unemployment",
  "Youth unemployment"
)


# Construct the latest-available cross-section

covariates <- matrix(
  NA_real_,
  nrow = nrow(dat),
  ncol = length(covariate_names),
  dimnames = list(
    dat$iso2c,
    covariate_names
  )
)


latest_wdi <- wdi |>
  select(iso2c,year,all_of(covariate_names)) |>
  pivot_longer(
    cols = all_of(covariate_names),
    names_to = "variable",
    values_to = "value"
  ) |>
  filter(!is.na(value)) |>
  group_by(iso2c,variable) |>
  slice_max(
    year,
    n = 1,
    with_ties = FALSE
  ) |>
  ungroup() |>
  select(iso2c,variable,value) |>
  pivot_wider(names_from = variable,values_from = value)


covariates_df <- data.frame(iso2c = dat$iso2c) |>
  left_join(
    latest_wdi,
    by = "iso2c"
  )

covariates <- as.matrix(
  covariates_df[
    covariate_names
  ]
)

rownames(covariates) <-
  covariates_df$iso2c


# Add intergenerational income persistence
# Higher IGE means lower intergenerational mobility

covariates <- cbind(
  covariates,
  IGE = as.numeric(dat$IGE)
)

covariate_names <- c(
  covariate_names,
  "IGE"
)

panel_names <- c(
  panel_names,
  "Intergenerational\nincome persistence"
)


# Correlations and HC1 confidence intervals

# Make 3 matrices for the CIs and results
correlations <- matrix(
  NA_real_,
  nrow = length(rho_alt),
  ncol = length(covariate_names),
  dimnames = list(
    rho_alt,
    covariate_names
  )
)

ci_lower <- correlations
ci_upper <- correlations

# loop through all the covariates 

# to inspect the models we introduce an audit vector

audit_models <- list()

for (v in seq_along(covariate_names)) {

  for (j in seq_along(rho_alt)) {

    ok <- is.finite(covariates[, v]) &
      is.finite(eta_alt[, j])

    if (sum(ok) < 3) {
      next
    }

    x <- covariates[ok, v]
    y <- eta_alt[ok, j]

    model <- lm(
      scale(y) ~ scale(x)
    )

    audit_models[[paste(covariate_names[v],rho_alt[j],sep = "_")]] <- model
    
    robust_result <- coeftest(
      model,
      vcov. = vcovHC(
        model,
        type = "HC1"
      )
    )

    correlations[j, v] <-
      robust_result[2, 1]

    ci_lower[j, v] <-
      robust_result[2, 1] -
      1.96 * robust_result[2, 2]

    ci_upper[j, v] <-
      robust_result[2, 1] +
      1.96 * robust_result[2, 2]
  }
}

```

## plots

```{r}

par(
  mfrow = c(4, 4),
  oma = c(4, 4, 5, 1),
  mar = c(2, 2.5, 2.5, 1)
)

for (v in seq_along(covariate_names)) {

  plot(
    100 * rho_alt,
    correlations[, v],
    type = "n",
    xaxt = "n",
    xlab = "",
    ylab = "",
    ylim = range(
      ci_lower[, v],
      ci_upper[, v],
      na.rm = TRUE
    ),
    main = panel_names[v]
  )

  polygon(
    c(
      100 * rho_alt,
      rev(100 * rho_alt)
    ),
    c(
      ci_lower[, v],
      rev(ci_upper[, v])
    ),
    col = adjustcolor(
      "#1F4E79",
      alpha.f = 0.15
    ),
    border = NA
  )

  grid(
    col = "grey90"
  )

  abline(
    h = 0,
    col = "grey60"
  )

  lines(
    100 * rho_alt,
    correlations[, v],
    col = "#1F4E79",
    lwd = 2
  )

  axis(
    1,
    at = c(
      3,
      10,
      20,
      30,
      40
    ),
    labels = paste0(
      c(
        3,
        10,
        20,
        30,
        40
      ),
      "%"
    )
  )

  sign_change <- which(
    is.finite(
      correlations[
        -nrow(correlations),
        v
      ]
    ) &
      is.finite(
        correlations[
          -1,
          v
        ]
      ) &
      correlations[
        -nrow(correlations),
        v
      ] *
      correlations[
        -1,
        v
      ] < 0
  )[1]

  if (!is.na(sign_change)) {

    abline(
      v = 100 * rho_alt[sign_change],
      col = "red",
      lwd = 2
    )
  }
}

mtext(
  "Subjective discount rate",
  side = 1,
  outer = TRUE,
  line = 2.2
)

mtext(
  expression(
    "Correlation with " * eta[j]
  ),
  side = 2,
  outer = TRUE,
  line = 2.2
)

mtext(
  "Which correlations survive the discounting assumption",
  side = 3,
  outer = TRUE,
  line = 2.5,
  cex = 1.2
)

mtext(
  paste(
    "Shaded bands show 95% HC1 confidence intervals;",
    "red line marks a sign change"
  ),
  side = 3,
  outer = TRUE,
  line = 1.2,
  cex = 0.85
)

par(
  mfrow = c(1, 1)
)

```

