---
title: "Early subjective-discount calculations — reproducible reconstruction"
output:
  html_document:
    df_print: paged
---

This file reconstructs the early full-contract subjective-discount calculations discussed in the chat.

Important distinction. `simulations_PVs_audit_v1_original.Rmd` is the older audit file and uses the simplified common-threshold loan model. The 3% to 30% subjective-discount sweep was not originally written into that RMD. This file makes that later calculation explicit and reproducible.

The reconstruction uses:

- `wages.xlsx`, sheet `main_df`
- `betas_saturated_model.xlsx`, sheet `main`
- the attached official GOV.UK 2025-26 Plan 2 overseas-threshold table
- World Bank indicator `NY.GDP.PCAP.PP.KD`, year 2022, only for the optional GDP-gradient section

The early calculation assumes the 2025-26 country thresholds remain fixed over the simulated 30-year horizon. It also assumes constant RPI of 4.3%, nominal wage growth of 2%, and no application-to-first-repayment lag. Those are calibration choices, not estimates.

```{r setup, message=FALSE, warning=FALSE}
library(readxl)
library(dplyr)
library(tidyr)
library(sandwich)
library(lmtest)

L <- 9250
Ppost <- 22200
alpha <- 0.09
wage_growth <- 0.02
rpi <- 0.043
interest_spread <- 0.03
horizon <- 30

# Eurostat wages are already expressed in euros for every country.
# Therefore the relevant conversion is EUR -> GBP, not each country's local-currency rate.
eur_gbp <- 0.848248
```

# 1. Official 2025-26 Plan 2 overseas thresholds

These lower and upper thresholds are transcribed from the attached GOV.UK file.

```{r thresholds}
thresholds <- tribble(
  ~Country,      ~iso2c, ~lower, ~upper,
  "Austria",     "AT", 28470, 51245,
  "Belgium",     "BE", 28470, 51245,
  "Bulgaria",    "BG", 11390, 20500,
  "Croatia",     "HR", 17085, 30750,
  "Cyprus",      "CY", 22780, 41000,
  "Czechia",     "CZ", 17085, 30750,
  "Denmark",     "DK", 28470, 51245,
  "Estonia",     "EE", 22780, 41000,
  "Finland",     "FI", 28470, 51245,
  "France",      "FR", 22780, 41000,
  "Germany",     "DE", 28470, 51245,
  "Greece",      "GR", 17085, 30750,
  "Hungary",     "HU", 17085, 30750,
  "Ireland",     "IE", 28470, 51245,
  "Italy",       "IT", 22780, 41000,
  "Latvia",      "LV", 17085, 30750,
  "Lithuania",   "LT", 17085, 30750,
  "Luxembourg",  "LU", 28470, 51245,
  "Malta",       "MT", 22780, 41000,
  "Netherlands", "NL", 28470, 51245,
  "Poland",      "PL", 17085, 30750,
  "Portugal",    "PT", 17085, 30750,
  "Romania",     "RO", 11390, 20500,
  "Slovakia",    "SK", 17085, 30750,
  "Slovenia",    "SI", 22780, 41000,
  "Spain",       "ES", 22780, 41000,
  "Sweden",      "SE", 28470, 51245
)
```

# 2. Input data

```{r inputs}
wages <- read_excel("wages.xlsx", sheet = "main_df") %>%
  inner_join(thresholds, by = "Country") %>%
  mutate(wage_gbp = Wage_EURO * eur_gbp)

betas <- read_excel("betas_saturated_model.xlsx", sheet = "main") %>%
  mutate(
    Estimate_TWFE = as.numeric(Estimate_TWFE),
    Estimate_controls = as.numeric(Estimate_controls),
    Estimate_Pooled = as.numeric(Estimate_Pooled)
  )

dat <- wages %>%
  left_join(
    betas[, c(
      "Country",
      "Estimate_TWFE",
      "Estimate_controls",
      "Estimate_Pooled"
    )],
    by = "Country"
  ) %>%
  filter(!is.na(Estimate_TWFE))

stopifnot(nrow(dat) == 25)
```

# 3. Full-contract repayment stream

The loan balance earns contractual interest. The interest rate is RPI plus an income-dependent loading between 0 and 3 percentage points.

```{r loan-engine}
loan_cashflows <- function(
    y0,
    lower,
    upper,
    L = 9250,
    alpha = 0.09,
    wage_growth = 0.02,
    rpi = 0.043,
    interest_spread = 0.03,
    horizon = 30) {

  balance <- L
  out <- vector("list", horizon)

  for (t in seq_len(horizon)) {

    earnings <- y0 * (1 + wage_growth)^(t - 1)

    loading <- if (earnings <= lower) {
      0
    } else if (earnings >= upper) {
      1
    } else {
      (earnings - lower) / (upper - lower)
    }

    contract_rate <- rpi + interest_spread * loading

    balance_after_interest <- balance * (1 + contract_rate)

    scheduled_repayment <- alpha * max(0, earnings - lower)
    payment <- min(scheduled_repayment, balance_after_interest)

    end_balance <- balance_after_interest - payment

    out[[t]] <- data.frame(
      t = t,
      earnings = earnings,
      loading = loading,
      contract_rate = contract_rate,
      scheduled_repayment = scheduled_repayment,
      payment = payment,
      end_balance = end_balance
    )

    balance <- end_balance

    if (balance <= 1e-10) break
  }

  bind_rows(out)
}
```

# 4. Subjective present value

This is the key change. The contractual repayment stream is generated first. Then the same cash flows are discounted using a subjective discount rate `rho`.

The early calculation did not add a separate lag from the application date to the first repayment year. Therefore year 1 of the simulated repayment stream is discounted by one period.

```{r subjective-pv}
subjective_pv <- function(y0, lower, upper, rho) {

  cf <- loan_cashflows(
    y0 = y0,
    lower = lower,
    upper = upper,
    L = L,
    alpha = alpha,
    wage_growth = wage_growth,
    rpi = rpi,
    interest_spread = interest_spread,
    horizon = horizon
  )

  sum(cf$payment / (1 + rho)^cf$t)
}
```

# 5. Reproduce the subjective-discount sweep

```{r sweep}
rho_grid <- c(0.03, 0.05, 0.075, 0.10, 0.125, 0.15, 0.20, 0.25, 0.30)

results <- crossing(
  iso2c = dat$iso2c,
  rho = rho_grid
) %>%
  left_join(dat, by = "iso2c") %>%
  rowwise() %>%
  mutate(
    PV_subjective = subjective_pv(
      y0 = wage_gbp,
      lower = lower,
      upper = upper,
      rho = rho
    ),
    k_i = log(Ppost) - log(PV_subjective),
    eta_TWFE = Estimate_TWFE / k_i,
    eta_controls = Estimate_controls / k_i,
    eta_Pooled = Estimate_Pooled / k_i
  ) %>%
  ungroup()

sweep_summary <- results %>%
  group_by(rho) %>%
  summarise(
    n = n(),
    mean_PV = mean(PV_subjective),
    mean_eta_TWFE = mean(eta_TWFE),
    mean_eta_controls = mean(eta_controls),
    median_eta_controls = median(eta_controls),
    .groups = "drop"
  )

sweep_summary
```

For a basic audit, the mean subjective PV should be approximately:

```{r expected-pv-check}
expected <- tibble(
  rho = c(0.03, 0.05, 0.075, 0.10, 0.125, 0.15, 0.20, 0.25, 0.30),
  expected_mean_PV = c(
    11101.60, 9029.17, 7236.57, 5999.04, 5108.92,
    4444.80, 3528.03, 2927.91, 2504.41
  ),
  expected_mean_eta_controls = c(
    -1.6511, -1.2117, -0.9519, -0.8046, -0.7082,
    -0.6398, -0.5484, -0.4897, -0.4485
  )
)

check <- sweep_summary %>%
  select(rho, mean_PV, mean_eta_controls) %>%
  left_join(expected, by = "rho") %>%
  mutate(
    PV_difference = mean_PV - expected_mean_PV,
    eta_difference = mean_eta_controls - expected_mean_eta_controls
  )

check
```

# 6. Country elasticities at the rates discussed in the chat

```{r country-table}
country_table <- results %>%
  filter(rho %in% c(0.03, 0.10, 0.15, 0.20)) %>%
  select(Country, rho, PV_subjective, k_i, eta_controls) %>%
  arrange(Country, rho)

country_table

country_wide <- country_table %>%
  select(Country, rho, eta_controls) %>%
  mutate(rho = paste0(round(100 * rho), "%")) %>%
  pivot_wider(names_from = rho, values_from = eta_controls)

country_wide
```

# 7. Inspect one country's repayment path

Change `inspect_country` and `inspect_rho` as needed.

```{r inspect-country}
inspect_country <- "Bulgaria"
inspect_rho <- 0.20

one <- dat %>% filter(Country == inspect_country)

cf_one <- loan_cashflows(
  y0 = one$wage_gbp,
  lower = one$lower,
  upper = one$upper,
  L = L,
  alpha = alpha,
  wage_growth = wage_growth,
  rpi = rpi,
  interest_spread = interest_spread,
  horizon = horizon
) %>%
  mutate(
    subjective_discount_factor = 1 / (1 + inspect_rho)^t,
    subjective_PV_contribution = payment * subjective_discount_factor
  )

cf_one
sum(cf_one$subjective_PV_contribution)
```

# 8. Optional GDP-gradient check

The early GDP-gradient calculation used 2022 World Bank PPP GDP per capita, indicator `NY.GDP.PCAP.PP.KD`.

This section requires internet access and the `WDI` package. World Bank historical series can be revised, so a fresh API pull may not reproduce an old regression to the last decimal unless a dated snapshot is saved.

```{r gdp-gradient, eval=FALSE}
library(WDI)

wb <- WDI(
  country = unique(dat$iso2c),
  indicator = c(GDP_PPP = "NY.GDP.PCAP.PP.KD"),
  start = 2022,
  end = 2022
) %>%
  select(iso2c, GDP_PPP) %>%
  mutate(log_gdp = log(GDP_PPP))

gradient_results <- results %>%
  left_join(wb, by = "iso2c") %>%
  filter(!is.na(log_gdp)) %>%
  group_by(rho) %>%
  group_modify(~{
    m <- lm(eta_controls ~ log_gdp, data = .x)
    ct <- coeftest(m, vcov. = vcovHC(m, type = "HC1"))

    tibble(
      n = nrow(.x),
      correlation = cor(.x$eta_controls, .x$log_gdp),
      slope = unname(ct["log_gdp", "Estimate"]),
      robust_se = unname(ct["log_gdp", "Std. Error"]),
      p_value = unname(ct["log_gdp", "Pr(>|t|)"])
    )
  }) %>%
  ungroup()

gradient_results
```

# 9. Sensitivity that should be added before publication

The early calculation above is intentionally reproduced as it was conceptually discussed. It is not the final behavioral specification.

A publication version should separately test:

1. a lag from application to the first repayment year
2. time-varying overseas thresholds instead of holding the 2025-26 table fixed
3. alternative RPI paths
4. post-Brexit tuition payment timing
5. subjective discount rates justified independently from the literature
6. quasi-hyperbolic discounting as an alternative to constant exponential discounting
