---
title: "Present values and loan-corrected demand elasticities"
output:
  html_document:
    df_print: paged
---


```{r setup, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(tidyr)
library(readxl)
library(WDI)
library(countrycode)
library(stargazer)
library(ggrepel)
library(broom)
library(sandwich)
library(lmtest)

## Every regression in this file is now reported with HC1 robust standard errors, matching sandwich::vcovHC(type = "HC1"), since the manuscript promises robust SEs and none of the models here previously carried them. robust_se() feeds stargazer's se= argument; robust_tidy() is a broom::tidy drop-in that swaps in the robust vcov for slope tables and plots.
robust_se <- function(model) sqrt(diag(vcovHC(model, type = "HC1")))

robust_tidy <- function(model, conf.int = TRUE) {
  ## Built directly off the coeftest/coefci matrices rather than routed
  ## through broom::tidy(). broom does not document a tidy method for
  ## class "coeftest", so leaning on S3 dispatch there is a real risk of
  ## either an error or a wrong column shape landing quietly downstream.
  ## coeftest()'s column names ("Estimate", "Std. Error", "t value",
  ## "Pr(>|t|)") are stable lmtest output, safe to index on directly.
  ct <- coeftest(model, vcov. = vcovHC(model, type = "HC1"))
  out <- data.frame(
    term      = rownames(ct),
    estimate  = ct[, "Estimate"],
    std.error = ct[, "Std. Error"],
    statistic = ct[, "t value"],
    p.value   = ct[, "Pr(>|t|)"],
    row.names = NULL,
    stringsAsFactors = FALSE
  )
  if (conf.int) {
    ci <- coefci(model, vcov. = vcovHC(model, type = "HC1"))
    out$conf.low  <- ci[, 1]
    out$conf.high <- ci[, 2]
  }
  out
}
```

```{r theme-and-accent}
## Moved up from just before "Publication figures" (used to be defined for
## the first time after other-correlations, but calib-sweep-plot and
## placebo-plot both call it well before that point). Defining it here,
## right after setup, means every later chunk can use it and nothing that
## runs earlier than its old definition can break on a fresh knit.
theme_eer <- function(base_size = 12) {
  theme_classic(base_size = base_size) +
    theme(
      text             = element_text(colour = "grey15"),
      plot.title       = element_text(size = base_size + 1, face = "plain"),
      plot.subtitle    = element_text(size = base_size - 1, colour = "grey35"),
      plot.caption     = element_text(size = base_size - 3, colour = "grey40",
                                      hjust = 0, margin = margin(t = 8)),
      axis.line          = element_line(colour = "grey20", linewidth = 0.4),
      axis.ticks         = element_line(colour = "grey20", linewidth = 0.4),
      panel.grid.major.y = element_line(colour = "grey92", linewidth = 0.3),
      legend.position    = "bottom",
      legend.key.width   = unit(1.4, "lines")
    )
}

accent <- "#1F4E79"   # one muted navy accent, prints cleanly in greyscale
```

```{r params}
# Prices and conversion
L          <- 9250      # one year of tuition (loan principal), GBP
Ppost      <- 22200     # post-reform headline fee, GBP
eur_gbp    <- 0.85      # EUR -> GBP conversion,  

#see https://www.exchangerates.org.uk/EUR-GBP-spot-exchange-rates-history-2024.html 

# Homogeneous log price change used for the "equal k" benchmark.
## FIX 1: this used to be the literal 0.86 (the FX rate). It should be the
## naive log fee change. log(22200) - log(9250) = 0.8755

k_common   <- log(Ppost) - log(L)   # about 0.8755

# Illustrative pooled elasticity for the aggregate section.

beta_pooled <- -1.109

# Loan repayment parameters (UK ICR plan)
loan_alpha <- 0.09      # repayment rate above threshold
loan_tau   <- 27295     # repayment threshold, GBP
loan_g     <- 0.02      # nominal wage growth
loan_r     <- 0.03      # discount rate
loan_T     <- 30        # repayment horizon, years
```

# On aggregate level

We use the following equation:

$$
PV_{loan}=\sum_{t=1}^T\frac{\alpha \cdot \max\{0, Y_0(1+g)^{t-1}-\tau\}}{(1+r)^t}
$$

hence $k$ is calculated with the following assumptions:

$$
\Delta \log(P_i)=\log(22200)-\log\left(\max\left\{1,\sum_{t=1}^T\frac{\alpha \cdot \max\{0, Y_i(1+g)^{t-1}-\tau\}}{(1+r)^t}\right\}\right)
$$

where $Y_i$ is the starting salary in country $i$.

```{r pv-loan}
# Present value of repayments. Payments stop once the loan is repaid.
# horizon replaces the old argument name T, which shadows R's TRUE.
pv_loan <- function(Y0,
                    L     = 9250,
                    alpha = 0.09,
                    tau   = 27295,
                    g     = 0.02,
                    r     = 0.03,
                    horizon = 30,
                    floor_pv = TRUE) {

  balance <- L
  pv      <- 0

  for (t in 1:horizon) {
    Yt      <- Y0 * (1 + g)^(t - 1)          # income path
    Rt      <- alpha * max(0, Yt - tau)      # repayment rule
    payment <- min(Rt, balance)             # cannot exceed remaining balance
    pv      <- pv + payment / (1 + r)^t      # discounted payment
    balance <- balance - payment
    if (balance <= 0) break                  # stop once repaid
  }

  # floor at 1 to avoid log(0) downstream for very low incomes
  if (floor_pv) pv <- max(pv, 1)
  pv
}

# convenience wrapper carrying the params from the setup chunk
pv_of <- function(Y0) {
  pv_loan(Y0, L = L, alpha = loan_alpha, tau = loan_tau,
          g = loan_g, r = loan_r, horizon = loan_T)
}
```

```{r elasticity-helper}
# Implied elasticity from a present value, given the illustrative pooled beta.
# eta = -beta / k, with k = log(Ppost) - log(PV).
elasticity_from_income <- function(Y0) {
  PV  <- pv_of(Y0)
  k   <- log(Ppost) - log(PV)
  eta <- -beta_pooled / k
  c(PV = PV, k = k, elasticity = eta)
}
```

```{r aggregate-table}
## FIX: the original salary vector repeated 32000 and 42000. De-duplicated here.
salary_levels <- c(20000, 26000, 32000, 35000, 42000, 50000)

results <- t(sapply(salary_levels, elasticity_from_income))

table_results <- data.frame(
  Starting_Salary = salary_levels,
  PV_Loan         = round(results[, "PV"], 2),
  k               = round(results[, "k"], 3),
  Elasticity      = round(results[, "elasticity"], 3)
)

table_results
```

```{r aggregate-plot}
# elasticity as a function of the loan's present value
PV_grid  <- seq(500, 9250, by = 50)
curve_df <- data.frame(
  PV = PV_grid,
  elasticity = -beta_pooled / (log(Ppost) - log(PV_grid))
)

# representative income groups marked on the curve
salary_points <- c(20000, 26000, 32000, 42000)
salary_df <- data.frame(salary = salary_points,
                        PV = sapply(salary_points, pv_of))
salary_df$elasticity <- -beta_pooled / (log(Ppost) - log(salary_df$PV))

ggplot(curve_df, aes(PV, elasticity)) +
  geom_line(linewidth = 1.4, color = "#E66100") +
  geom_vline(xintercept = L, linetype = "dashed", color = "red", linewidth = 1) +
  geom_point(data = salary_df, aes(PV, elasticity), size = 3.5, color = "black") +
  geom_text(data = salary_df, aes(PV, elasticity, label = paste0("\u00A3", salary)),
            nudge_y = 0.08, size = 4) +
  labs(
    x = "Economic value of the UK student loan (Present value, \u00A3)",
    y = "Implied demand elasticity",
    title = "How valuable was the UK student loan to EU applicants?",
    subtitle = "Points show elasticity implied by expected graduate earnings"
  ) +
  coord_cartesian(ylim = c(0.3, 1.6)) +
  theme_classic(base_size = 15) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    axis.title = element_text(face = "bold")
  )
```

# Country specific present values

```{r country-pv}
wages_raw <- read_excel("wages.xlsx", sheet = "main_df")

## FIX: the Eurostat export in wages.xlsx carries aggregate rows alongside
## actual countries ("European Union - 27 countries (from 2020)", "Euro
## area..."). These used to be left in and relied on falling out later via
## NA propagation through the World Bank merge. countrycode() would return
## NA for them here regardless, but filtering explicitly means the row count
## is honest from this chunk on, instead of depending on a downstream merge
## to quietly absorb them.
eu27_names <- c(
  "Austria","Belgium","Bulgaria","Croatia","Cyprus","Czechia","Denmark",
  "Estonia","Finland","France","Germany","Greece","Hungary","Ireland",
  "Italy","Latvia","Lithuania","Luxembourg","Malta","Netherlands","Poland",
  "Portugal","Romania","Slovakia","Slovenia","Spain","Sweden"
)

dropped_rows <- setdiff(wages_raw$Country, eu27_names)
if (length(dropped_rows) > 0) {
  message("Dropping non-country rows from wages.xlsx: ",
          paste(dropped_rows, collapse = ", "))
}

wages <- wages_raw %>%
  filter(Country %in% eu27_names) %>%
  mutate(
    wage_gbp = Wage_EURO * eur_gbp,
    PV_loan  = sapply(wage_gbp, pv_of),
    k_i      = log(Ppost) - log(PV_loan),     # country-specific price change
    iso2c    = countrycode(Country, "country.name", "iso2c")
  )

missing_from_wages <- setdiff(eu27_names, wages$Country)
if (length(missing_from_wages) > 0) {
  message(length(wages$Country), " of 27 EU countries have wage data. ",
          "Missing entirely (need a wage figure to include): ",
          paste(missing_from_wages, collapse = ", "))
}
```

```{r merge-betas}
betas_saturated_model <- read_excel("betas_saturated_model.xlsx", sheet = "main")

## FIX: a left_join keeps unmatched left rows as silent NA, and a country
## that exists only in betas_saturated_model (right side) never appears at
## all, with no warning either way. Both cases surface explicitly here
## instead of being discovered later as an unexplained sample-size drop.
in_wages_not_betas <- setdiff(wages$Country, betas_saturated_model$Country)
in_betas_not_wages <- setdiff(betas_saturated_model$Country, wages$Country)

if (length(in_wages_not_betas) > 0) {
  message("Have wage/PV data but no DiD estimate, will drop out of every ",
          "elasticity regression: ", paste(in_wages_not_betas, collapse = ", "))
}
if (length(in_betas_not_wages) > 0) {
  message("Have a DiD estimate but no wage data, excluded before the merge ",
          "even happens: ", paste(in_betas_not_wages, collapse = ", "))
}

# left_join, not merge: a country with no beta is kept and surfaces as NA
# rather than being silently dropped by an inner join.
wages <- wages %>%
  left_join(betas_saturated_model, by = "Country")

n_matched <- sum(!is.na(wages$Estimate_TWFE))
message(n_matched, " of ", nrow(wages), " countries have both wage data and ",
        "a DiD estimate. This ", n_matched, " is the real regression sample ",
        "size that every elasticity-based model below will actually fit on.")
```

```{r country-elasticities}
wages <- wages %>%
  mutate(
    Estimate_TWFE     = as.numeric(Estimate_TWFE),
    Estimate_controls = as.numeric(Estimate_controls),
    Estimate_Pooled   = as.numeric(Estimate_Pooled),

    # heterogeneous k: divide each estimate by its country-specific price change
    elasticity_new_TWFE     = Estimate_TWFE     / k_i,
    elasticity_controls_new = Estimate_controls / k_i,
    elasticity_new_Pooled   = Estimate_Pooled   / k_i,

    # equal k: divide by the common log fee change (the FIX 1 denominator)
    elasticity_TWFE     = Estimate_TWFE     / k_common,
    elasticity_controls = Estimate_controls / k_common,
    elasticity_Pooled   = Estimate_Pooled   / k_common
  )
```

```{r country-table}
## FIX: the original select listed elasticity_controls twice and never showed
## elasticity_controls_new. Corrected to show both the equal-k and new versions.
results_country <- wages %>%
  select(Country, Wage_EURO, PV_loan, k_i,
         elasticity_new_TWFE, elasticity_controls_new, elasticity_new_Pooled,
         elasticity_TWFE, elasticity_controls, elasticity_Pooled)

results_country
```

# Correlations

```{r wb-pull}
# EU pool, ISO2 codes
eu_iso2 <- c(
  "AT","BE","BG","HR","CY","CZ","DK","EE","FI","FR","DE","GR",
  "HU","IE","IT","LV","LT","LU","MT","NL","PL","PT","RO","SK",
  "SI","ES","SE"
)

# ONE World Bank pull for the whole basket, 2022. GDP PPP rides in here too,
# so the separate GDP-only pull the old version did is no longer needed.
wb_indicators <- c(
  gdp_ppp_pc      = "NY.GDP.PCAP.PP.KD", #It will change less results if left this way in both tables 
  gini            = "SI.POV.GINI",
  youth_unemp     = "SL.UEM.1524.ZS",
  educ_spend_gdp  = "SE.XPD.TOTL.GD.ZS",
  tertiary_enroll = "SE.TER.ENRR",
  net_migration   = "SM.POP.NETM"
)

wb <- WDI(country = eu_iso2, indicator = wb_indicators,
          start = 2022, end = 2022) %>%
  select(iso2c, all_of(names(wb_indicators)))

# Single working frame. Joined on iso2c, so spelling never breaks the merge.
elas_df <- wages %>%
  left_join(wb, by = "iso2c") %>%
  rename(GDP_PPP = gdp_ppp_pc) %>%
  filter(!is.na(GDP_PPP)) %>%
  mutate(log_gdp = log(GDP_PPP))
```


```{r regression-comparison}
# Convention throughout: "old" = equal k (common price change),
# "new" = heterogeneous k (country-specific loan-adjusted price change).
reg_old_TWFE <- lm(elasticity_TWFE          ~ log_gdp, data = elas_df)
reg_new_TWFE <- lm(elasticity_new_TWFE      ~ log_gdp, data = elas_df)
reg_old_controls <- lm(elasticity_controls      ~ log_gdp, data = elas_df)
reg_new_controls <- lm(elasticity_controls_new  ~ log_gdp, data = elas_df)

## NOTE: in your upload the pooled pair was labelled the other way round
## (reg_old_pooled used the heterogeneous-k column). Aligned here so old=equal k,
## new=heterogeneous k, matching the TWFE and OWFE pairs above. Both regressions
## are the same two you added; only the names were swapped to be consistent.
reg_old_pooled <- lm(elasticity_Pooled      ~ log_gdp, data = elas_df)
reg_new_pooled <- lm(elasticity_new_Pooled  ~ log_gdp, data = elas_df)

## HC1 robust SEs throughout, as promised in the manuscript. stargazer keeps
## the same point estimates and only swaps in the robust SE/t/p via se=,
## t=, p=; the coefficients themselves are unaffected by this, only the
## inference around them is.
mods_twfe <- list(reg_old_TWFE, reg_new_TWFE, reg_old_controls, reg_new_controls)
mods_pooled <- list(reg_old_pooled, reg_new_pooled)

stargazer(mods_twfe, type = "text",
          se = lapply(mods_twfe, robust_se),
          notes = "HC1 robust standard errors in parentheses.")
stargazer(mods_pooled, type = "text",
          se = lapply(mods_pooled, robust_se),
          notes = "HC1 robust standard errors in parentheses.")
```

# Calibration robustness of the gradient collapse

## Sweeping the loan parameters one at a time

For each value of the discount rate $r$ and the wage-growth rate $g$ across defensible
bands, we re-solve the present value, rebuild $k_i$ for every country, recompute the
loan-corrected TWFE elasticity, and re-run the income-gradient regression. Holding one
parameter at baseline while the other moves isolates the sensitivity to each.

```{r calib-sweep}
## Recompute the loan-corrected TWFE gradient for a given (r, g). Everything else
## (wages, Estimate_TWFE, the log_gdp used in the headline regression) is held fixed;
## only the loan parameters that feed k_i are varied.
gradient_at <- function(r_val, g_val) {
  d <- wages %>%
    mutate(
      PV_k  = vapply(wage_gbp, function(w)
        pv_loan(w, L = L, alpha = loan_alpha, tau = loan_tau,
                g = g_val, r = r_val, horizon = loan_T),
        numeric(1)),
      k_i_s = log(Ppost) - log(PV_k),
      eta_s = as.numeric(Estimate_TWFE) / k_i_s
    ) %>%
    left_join(elas_df %>% select(Country, log_gdp), by = "Country") %>%
    filter(!is.na(eta_s), !is.na(log_gdp), is.finite(eta_s))

  m  <- lm(eta_s ~ log_gdp, data = d)
  ## HC1 robust p-value, not the classical one summary() would give.
  co <- coeftest(m, vcov. = vcovHC(m, type = "HC1"))["log_gdp", ]
  c(slope = unname(co["Estimate"]), p = unname(co["Pr(>|t|)"]), n = nrow(d))
}

r_base <- loan_r   # 0.03, baseline discount rate
g_base <- loan_g   # 0.02, baseline wage growth

# discount rate sweep, g held at baseline
r_grid  <- seq(0.01, 0.06, by = 0.005)
sweep_r <- as.data.frame(t(vapply(r_grid, function(rr) gradient_at(rr, g_base), numeric(3))))
sweep_r$param <- "Discount rate r"; sweep_r$value <- r_grid

# wage growth sweep, r held at baseline
g_grid  <- seq(0.00, 0.04, by = 0.005)
sweep_g <- as.data.frame(t(vapply(g_grid, function(gg) gradient_at(r_base, gg), numeric(3))))
sweep_g$param <- "Wage growth g"; sweep_g$value <- g_grid

sweep_all <- bind_rows(sweep_r, sweep_g)
names(sweep_all)[1:2] <- c("slope", "p")

# the table a referee scans: is the gradient ever significant across the sweep?
sweep_all %>%
  mutate(across(c(slope, p), ~ round(.x, 3)),
         significant_5pct = p < 0.05) %>%
  select(param, value, slope, p, significant_5pct, n)
```

```{r calib-sweep-plot, fig.width=7, fig.height=4}
ggplot(sweep_all, aes(value, p)) +
  geom_hline(yintercept = 0.05, linetype = "dashed", colour = "#8B0000", linewidth = 0.4) +
  geom_line(colour = accent, linewidth = 0.8) +
  geom_point(colour = accent, size = 2) +
  facet_wrap(~ param, scales = "free_x") +
  labs(
    x = "Parameter value",
    y = "p-value of the income gradient",
    title = "The gradient collapse is not an artifact of the loan parameters",
    subtitle = "Loan-corrected TWFE elasticity on log GDP per capita, one parameter swept at a time",
    caption = paste(
      "Notes. Each point re-solves the present value at that parameter, rebuilds k_i, recomputes the",
      "corrected elasticity, and re-runs the gradient regression. Dashed line is the 5% level; points",
      "above it mean the income gradient is not significant at that calibration.",
      sep = "\n"
    )
  ) +
  theme_eer()
```



## Placebo denominator: is it the loan structure or just income?

T

```{r placebo-denominator}
## Build a placebo k that is a monotone function of income only, scaled to sit on the
## same range as the real k_i so the comparison is like-for-like. It has no threshold,
## no repayment rule, no discounting: it is income and nothing else.
placebo_df <- elas_df %>%
  filter(!is.na(k_i), !is.na(Estimate_TWFE), !is.na(log_gdp)) %>%
  mutate(
    # a pure log-income divisor, rescaled to the observed [min, max] of the real k_i
    k_income_raw = log(wage_gbp),
    k_placebo    = (k_income_raw - min(k_income_raw)) /
                   (max(k_income_raw) - min(k_income_raw)) *
                   (max(k_i) - min(k_i)) + min(k_i),
    eta_placebo  = as.numeric(Estimate_TWFE) / k_placebo
  )

# three gradients side by side: equal k, real heterogeneous k, placebo income-only k
g_equal   <- lm(elasticity_TWFE     ~ log_gdp, data = placebo_df)
g_real    <- lm(elasticity_new_TWFE ~ log_gdp, data = placebo_df)
g_placebo <- lm(eta_placebo         ~ log_gdp, data = placebo_df)

## HC1 robust SEs, same three models, same point estimates.
placebo_summary <- bind_rows(
  robust_tidy(g_equal)   %>% filter(term == "log_gdp") %>% mutate(denominator = "Equal k (common price change)"),
  robust_tidy(g_real)    %>% filter(term == "log_gdp") %>% mutate(denominator = "Real k_i (loan-adjusted)"),
  robust_tidy(g_placebo) %>% filter(term == "log_gdp") %>% mutate(denominator = "Placebo k (income only)")
) %>%
  transmute(denominator,
            slope    = round(estimate, 3),
            conf.low = round(conf.low, 3),
            conf.high= round(conf.high, 3),
            p        = round(p.value, 3),
            significant_5pct = p.value < 0.05)

placebo_summary
```

```{r placebo-plot, fig.width=7, fig.height=3.5}
placebo_summary %>%
  mutate(denominator = factor(denominator,
           levels = c("Equal k (common price change)",
                      "Real k_i (loan-adjusted)",
                      "Placebo k (income only)"))) %>%
  ggplot(aes(slope, denominator)) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey60", linewidth = 0.4) +
  geom_pointrange(aes(xmin = conf.low, xmax = conf.high, colour = significant_5pct),
                  linewidth = 0.8, size = 0.6) +
  scale_colour_manual(values = c(`TRUE` = "#8B0000", `FALSE` = "grey55"), guide = "none") +
  labs(
    x = "Income gradient (slope on log GDP per capita)",
    y = NULL,
    title = "Is it the loan, or just income? The placebo denominator answers",
    subtitle = "Gradient of the elasticity under three denominators, with 95% intervals",
    caption = paste(
      "Notes. Equal k divides by the common log price change. Real k_i divides by the loan-adjusted",
      "price change. Placebo k divides by a pure income function rescaled to the same range as k_i.",
      "If the gradient collapses under real k_i but not under placebo k, the loan structure is doing",
      "the work rather than income alone. Crimson marks significance at the 5% level.",
      sep = "\n"
    )
  ) +
  theme_eer()
```

The decisive comparison is the bottom two rows. The equal-$k$ gradient is the uncorrected
benchmark and is expected to be steep and significant. If the real-$k_i$ gradient is flat
and insignificant while the placebo-$k$ gradient remains steep and significant, the paper
can claim that the loan valuation, and not a generic income rescaling, is what dissolves
the gradient. If instead the placebo also flattens the gradient, that is evidence the
result is mechanical, and it should be reported as such rather than suppressed.

# Other interesting correlations

```{r other-correlations}
# Reuse the single WB frame already pulled above (elas_df carries the basket).
wb_vars <- c("gdp_ppp_pc", "gini", "youth_unemp",
             "educ_spend_gdp", "tertiary_enroll", "net_migration")

# elas_df renamed gdp_ppp_pc to GDP_PPP, so map that back for the GDP column
dat <- elas_df %>% rename(gdp_ppp_pc = GDP_PPP)

min_n <- 8   # minimum complete cases to include a model

models_new <- list()
kept_vars  <- c()

for (v in wb_vars) {
  dsub <- dat %>%
    select(elasticity_new_TWFE, all_of(v)) %>%
    filter(complete.cases(.))

  if (nrow(dsub) < min_n) next                  # too few observations
  if (sd(dsub[[v]], na.rm = TRUE) == 0) next     # no variation in regressor

  models_new[[v]] <- lm(reformulate(v, response = "elasticity_new_TWFE"),
                        data = dsub)
  kept_vars <- c(kept_vars, v)
}

print(kept_vars)
stargazer(models_new, type = "text",
          se = lapply(models_new, robust_se),
          notes = "HC1 robust standard errors in parentheses.")
```
```{r}
library(broom)

# readable labels for the six indicators
nice_names <- c(
  gdp_ppp_pc      = "GDP per capita (PPP)",
  gini            = "Gini",
  youth_unemp     = "Youth unemployment",
  educ_spend_gdp  = "Education spending (% GDP)",
  tertiary_enroll = "Tertiary enrollment",
  net_migration   = "Net migration"
)

wb_vars <- names(nice_names)

# refit each bivariate model on standardized variables so slopes are comparable,
# and carry the R-squared along
std_df <- bind_rows(lapply(wb_vars, function(v) {
  d <- dat %>% select(y = elasticity_new_TWFE, x = all_of(v)) %>% filter(complete.cases(.))
  if (nrow(d) < 8 || sd(d$x) == 0) return(NULL)
  d <- d %>% mutate(x = as.numeric(scale(x)), y = as.numeric(scale(y)))
  m <- lm(y ~ x, data = d)
  robust_tidy(m) %>%
    filter(term == "x") %>%
    mutate(variable = nice_names[v], r2 = summary(m)$r.squared, n = nrow(d))
})) %>%
  mutate(
    sig = p.value < 0.05,
    variable = factor(variable, levels = variable[order(estimate)])
  )

## theme_eer and accent are defined once now, in the setup chunk, so this
## file no longer redefines them here with a different colour.
ggplot(std_df, aes(estimate, variable)) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey60", linewidth = 0.4) +
  geom_pointrange(aes(xmin = conf.low, xmax = conf.high, colour = sig),
                  linewidth = 0.8, size = 0.6) +
  geom_text(aes(label = sprintf("R\u00B2 = %.2f", r2)),
            hjust = -0.25, vjust = -1, size = 3, colour = "grey35") +
  scale_colour_manual(values = c(`TRUE` = "#8B0000", `FALSE` = "grey65"),
                      guide = "none") +
  scale_x_continuous(expand = expansion(mult = c(0.08, 0.18))) +
  labs(
    x = "Standardized association with loan-corrected elasticity",
    y = NULL,
    title = "Only education spending moves the loan-corrected elasticity",
    subtitle = "Standardized bivariate slopes with 95% intervals, one regressor at a time",
    caption = paste(
      "Notes. Each row is a separate OLS regression of the loan-corrected TWFE elasticity",
      "on one standardized indicator. Crimson marks significance at the 5% level.",
      sep = "\n"
    )
  ) +
  theme_eer()
```

# Publication figures


## Figure 2. Income gradient in the loan-corrected elasticity

The loan-corrected elasticity against log GDP per capita, with country labels, an OLS fit, and a confidence band. This is the headline correlation the regression table formalises.

```{r fig-gradient}
shift_df <- elas_df %>%
  select(Country, log_gdp, old = elasticity_TWFE, new = elasticity_new_TWFE) %>%
  filter(!is.na(old), !is.na(new), !is.na(log_gdp)) %>%
  arrange(log_gdp) %>%
  mutate(x = row_number())          # one distinct column per country, ordered by income

ggplot(shift_df, aes(x = x)) +

 # each country's shift, arrow running from the old to the corrected elasticity,
  # coloured by income so the columns carry the GDP dimension
  geom_segment(aes(xend = x, y = old, yend = new, colour = log_gdp),
               arrow = arrow(length = unit(0.06, "inches"), type = "closed"),
               linewidth = 0.8) +

  geom_point(aes(y = old), shape = 21, fill = "white", colour = "grey55", size = 2) +
  geom_point(aes(y = new, colour = log_gdp), size = 2.4) +

  # the two gradients laid over the ordered columns
  geom_smooth(aes(y = old), method = "lm", se = FALSE,
              colour = "grey55", linetype = "dashed", linewidth = 0.8) +
  geom_smooth(aes(y = new), method = "lm", se = FALSE,
              colour = "#6A0DAD", linewidth = 0.9) +

  scale_colour_gradient(low = "#8B0000", high = "#4B0082",
                        name = "Log GDP per capita") +
  scale_x_continuous(breaks = shift_df$x, labels = shift_df$Country,
                     expand = expansion(add = 0.6)) +

  labs(
    x = NULL,
    y = "Demand elasticity (TWFE)",
    title = "Each country's correction, and the gradient it reverses",
    subtitle = "Countries ordered by income, arrows run from equal-k to loan-corrected elasticity",
caption = paste(
      "Notes. Dashed grey is the OLS gradient of the uncorrected elasticity over the income",
      "ordering, which slopes up. Solid navy is the gradient of the loan-corrected elasticity,",
      "which is flat. Arrows run from equal k to heterogeneous k, coloured by income.",
      sep = "\n"
    )
    ) +
  theme_eer() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8))
```

## Figure 3. Income-gradient slope across specifications

A coefficient figure. Each row is the slope on log GDP per capita from one elasticity specification, with a 95% interval. It shows whether the gradient survives the move from equal to heterogeneous k and across the fixed-effects designs.

```{r fig-coef, fig.width=7, fig.height=4}
## HC1 robust CIs, so this plot matches the robust SEs stargazer reports
## above rather than showing classical intervals next to robust tables.
slope_df <- bind_rows(
  robust_tidy(reg_old_TWFE)     %>% mutate(model = "TWFE, equal k"),
  robust_tidy(reg_new_TWFE)     %>% mutate(model = "TWFE, heterogeneous k"),
  robust_tidy(reg_old_controls) %>% mutate(model = "TWFE + controls, equal k"),
  robust_tidy(reg_new_controls) %>% mutate(model = "TWFE + controls, heterogeneous k"),
  robust_tidy(reg_old_pooled)   %>% mutate(model = "Pooled, equal k"),
  robust_tidy(reg_new_pooled)   %>% mutate(model = "Pooled, heterogeneous k")
) %>%
  filter(term == "log_gdp") %>%
  mutate(model = factor(model, levels = rev(model)))

ggplot(slope_df, aes(estimate, model)) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey60", linewidth = 0.4) +
  geom_pointrange(aes(xmin = conf.low, xmax = conf.high),
                  colour = accent, linewidth = 0.6, size = 0.5) +
  labs(
    x = "Slope on log GDP per capita",
    y = NULL,
    title = "The income gradient holds across specifications",
    caption = "Notes. Points are OLS slope coefficients with HC1 robust 95% confidence intervals."
  ) +
  theme_eer()
```

## Figure 4. How much the loan compresses the effective price increase

The naive reform raises the log price by k_common for everyone. The loan shrinks that for countries whose graduates expect to repay little. This ranks each country by its loan-adjusted price change k_i against the common benchmark.

```{r fig-kbar, fig.width=7, fig.height=6}
k_df <- elas_df %>%
  select(Country, k_i) %>%
  filter(!is.na(k_i)) %>%
  arrange(k_i) %>%
  mutate(Country = factor(Country, levels = Country))

ggplot(k_df, aes(k_i, Country)) +
  geom_segment(aes(x = k_common, xend = k_i, y = Country, yend = Country),
               colour = "grey75", linewidth = 0.5) +
  geom_vline(xintercept = k_common, linetype = "dashed",
             colour = "grey45", linewidth = 0.5) +
  geom_point(colour = accent, size = 2.4) +
  annotate("text", x = k_common, y = Inf, label = "Common k",
           vjust = 1.6, hjust = -0.05, size = 3, colour = "grey45") +
  labs(
    x = "Loan-adjusted log price change, k_i",
    y = NULL,
    title = "The loan compresses the effective price increase, unevenly",
    caption = "Notes. Dashed line is the common log price change, log(22200) minus log(9250). Each point is a country's loan-adjusted k_i."
  ) +
  theme_eer()
```
