---
title: "Present values and loan-corrected demand elasticities"
output:
  html_document:
    df_print: paged
---

<!--
Rebuild notes

  All constants live in one params chunk and are referenced everywhere
  (Ppost, L, eur_gbp, beta_pooled, the loan parameters, and k_common).

  pv_loan is defined ONCE.

  The World Bank data is pulled ONCE for the EU pool and joined on iso2c,
  so country-name spelling never gates a merge.

  Object names are unified. The working elasticity frame is elas_df
  (it replaces the generic df). Joins use left_join, never merge.

  Two substantive fixes are flagged inline with the tag  ## FIX:
    1. The "equal k" elasticities divided by 0.86, which is the EUR-GBP
       rate eur_gbp, not a price change. The homogeneous price change is
       k_common = log(Ppost) - log(L) = log(22200) - log(9250) = 0.8755.
       The two numbers are close, which is why the bug was invisible.
       Set k_common <- 0.86 below if you truly want the old behaviour.
    2. beta appeared as -1.28 in the table and -1.265 in the plot. It is
       now one constant, beta_pooled. Change it in one place if needed.
-->

```{r setup, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(tidyr)
library(readxl)
library(WDI)
library(countrycode)
library(stargazer)
library(ggrepel)
library(broom)
```

```{r params}
# Prices and conversion
L          <- 9250      # one year of tuition (loan principal), GBP
Ppost      <- 22200     # post-reform headline fee, GBP
eur_gbp    <- 0.86      # EUR -> GBP conversion

# Homogeneous log price change used for the "equal k" benchmark.
## FIX 1: this used to be the literal 0.86 (the FX rate). It should be the
## naive log fee change. log(22200) - log(9250) = 0.8755, close to 0.86,
## which is exactly why the mix-up went unnoticed.
k_common   <- log(Ppost) - log(L)   # about 0.8755

# Illustrative pooled elasticity for the aggregate section.
## FIX 2: was -1.28 in the table and -1.265 in the plot. One value now.
beta_pooled <- -1.08

# Loan repayment parameters (UK ICR plan)
loan_alpha <- 0.09      # repayment rate above threshold
loan_tau   <- 27295     # repayment threshold, GBP
loan_g     <- 0.02      # nominal wage growth
loan_r     <- 0.03      # discount rate
loan_T     <- 30        # repayment horizon, years
```

# On aggregate level

We use the following equation:

$$
PV_{loan}=\sum_{t=1}^T\frac{\alpha \cdot \max\{0, Y_0(1+g)^{t-1}-\tau\}}{(1+r)^t}
$$

hence $k$ is calculated with the following assumptions:

$$
\Delta \log(P_i)=\log(22200)-\log\left(\sum_{t=1}^T\frac{\alpha \cdot \max\{0, Y_i(1+g)^{t-1}-\tau\}}{(1+r)^t}\right)
$$

where $Y_i$ is the starting salary in country $i$.

```{r pv-loan}
# Present value of repayments. Payments stop once the loan is repaid.
# horizon replaces the old argument name T, which shadows R's TRUE.
pv_loan <- function(Y0,
                    L     = 9250,
                    alpha = 0.09,
                    tau   = 27295,
                    g     = 0.02,
                    r     = 0.03,
                    horizon = 30,
                    floor_pv = TRUE) {

  balance <- L
  pv      <- 0

  for (t in 1:horizon) {
    Yt      <- Y0 * (1 + g)^(t - 1)          # income path
    Rt      <- alpha * max(0, Yt - tau)      # repayment rule
    payment <- min(Rt, balance)             # cannot exceed remaining balance
    pv      <- pv + payment / (1 + r)^t      # discounted payment
    balance <- balance - payment
    if (balance <= 0) break                  # stop once repaid
  }

  # floor at 1 to avoid log(0) downstream for very low incomes
  if (floor_pv) pv <- max(pv, 1)
  pv
}

# convenience wrapper carrying the params from the setup chunk
pv_of <- function(Y0) {
  pv_loan(Y0, L = L, alpha = loan_alpha, tau = loan_tau,
          g = loan_g, r = loan_r, horizon = loan_T)
}
```

```{r elasticity-helper}
# Implied elasticity from a present value, given the illustrative pooled beta.
# eta = -beta / k, with k = log(Ppost) - log(PV).
elasticity_from_income <- function(Y0) {
  PV  <- pv_of(Y0)
  k   <- log(Ppost) - log(PV)
  eta <- -beta_pooled / k
  c(PV = PV, k = k, elasticity = eta)
}
```

```{r aggregate-table}
## FIX: the original salary vector repeated 32000 and 42000. De-duplicated here.
salary_levels <- c(20000, 26000, 32000, 35000, 42000, 50000)

results <- t(sapply(salary_levels, elasticity_from_income))

table_results <- data.frame(
  Starting_Salary = salary_levels,
  PV_Loan         = round(results[, "PV"], 2),
  k               = round(results[, "k"], 3),
  Elasticity      = round(results[, "elasticity"], 3)
)

table_results
```

```{r aggregate-plot}
# elasticity as a function of the loan's present value
PV_grid  <- seq(500, 9250, by = 50)
curve_df <- data.frame(
  PV = PV_grid,
  elasticity = -beta_pooled / (log(Ppost) - log(PV_grid))
)

# representative income groups marked on the curve
salary_points <- c(20000, 26000, 32000, 42000)
salary_df <- data.frame(salary = salary_points,
                        PV = sapply(salary_points, pv_of))
salary_df$elasticity <- -beta_pooled / (log(Ppost) - log(salary_df$PV))

ggplot(curve_df, aes(PV, elasticity)) +
  geom_line(linewidth = 1.4, color = "#E66100") +
  geom_vline(xintercept = L, linetype = "dashed", color = "red", linewidth = 1) +
  geom_point(data = salary_df, aes(PV, elasticity), size = 3.5, color = "black") +
  geom_text(data = salary_df, aes(PV, elasticity, label = paste0("\u00A3", salary)),
            nudge_y = 0.08, size = 4) +
  labs(
    x = "Economic value of the UK student loan (Present value, \u00A3)",
    y = "Implied demand elasticity",
    title = "How valuable was the UK student loan to EU applicants?",
    subtitle = "Points show elasticity implied by expected graduate earnings"
  ) +
  coord_cartesian(ylim = c(0.3, 1.6)) +
  theme_classic(base_size = 15) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    axis.title = element_text(face = "bold")
  )
```

# Country specific present values

```{r country-pv}
wages <- read_excel("wages.xlsx", sheet = "main_df")

wages <- wages %>%
  mutate(
    wage_gbp = Wage_EURO * eur_gbp,
    PV_loan  = sapply(wage_gbp, pv_of),
    k_i      = log(Ppost) - log(PV_loan),     # country-specific price change
    iso2c    = countrycode(Country, "country.name", "iso2c")
  )
```

```{r merge-betas}
betas_saturated_model <- read_excel("betas_saturated_model.xlsx")

# left_join, not merge: a country with no beta is kept and surfaces as NA
# rather than being silently dropped by an inner join.
wages <- wages %>%
  left_join(betas_saturated_model, by = "Country")
```

```{r country-elasticities}
wages <- wages %>%
  mutate(
    Estimate_TWFE     = as.numeric(Estimate_TWFE),
    Estimate_controls = as.numeric(Estimate_controls),
    Estimate_Pooled   = as.numeric(Estimate_Pooled),

    # heterogeneous k: divide each estimate by its country-specific price change
    elasticity_new_TWFE     = Estimate_TWFE     / k_i,
    elasticity_controls_new = Estimate_controls / k_i,
    elasticity_new_Pooled   = Estimate_Pooled   / k_i,

    # equal k: divide by the common log fee change (the FIX 1 denominator)
    elasticity_TWFE     = Estimate_TWFE     / k_common,
    elasticity_controls = Estimate_controls / k_common,
    elasticity_Pooled   = Estimate_Pooled   / k_common
  )
```

```{r country-table}
## FIX: the original select listed elasticity_controls twice and never showed
## elasticity_controls_new. Corrected to show both the equal-k and new versions.
results_country <- wages %>%
  select(Country, Wage_EURO, PV_loan, k_i,
         elasticity_new_TWFE, elasticity_controls_new, elasticity_new_Pooled,
         elasticity_TWFE, elasticity_controls, elasticity_Pooled)

results_country
```

# Correlations

```{r wb-pull}
# EU pool, ISO2 codes
eu_iso2 <- c(
  "AT","BE","BG","HR","CY","CZ","DK","EE","FI","FR","DE","GR",
  "HU","IE","IT","LV","LT","LU","MT","NL","PL","PT","RO","SK",
  "SI","ES","SE"
)

# ONE World Bank pull for the whole basket, 2022. GDP PPP rides in here too,
# so the separate GDP-only pull the old version did is no longer needed.
wb_indicators <- c(
  gdp_ppp_pc      = "NY.GDP.PCAP.PP.KD",
  gini            = "SI.POV.GINI",
  youth_unemp     = "SL.UEM.1524.ZS",
  educ_spend_gdp  = "SE.XPD.TOTL.GD.ZS",
  tertiary_enroll = "SE.TER.ENRR",
  net_migration   = "SM.POP.NETM"
)

wb <- WDI(country = eu_iso2, indicator = wb_indicators,
          start = 2022, end = 2022) %>%
  select(iso2c, all_of(names(wb_indicators)))

# Single working frame. Joined on iso2c, so spelling never breaks the merge.
elas_df <- wages %>%
  left_join(wb, by = "iso2c") %>%
  rename(GDP_PPP = gdp_ppp_pc) %>%
  filter(!is.na(GDP_PPP)) %>%
  mutate(log_gdp = log(GDP_PPP))
```

```{r gdp-gradient-plot, message=FALSE, warning=FALSE}
elas_long <- elas_df %>%
  pivot_longer(
    cols = c(elasticity_TWFE, elasticity_new_TWFE),
    names_to = "type",
    values_to = "elas"
  )

ggplot(elas_long, aes(x = log_gdp, y = elas, color = type)) +
  geom_point(size = 3) +
  geom_smooth(method = "lm", se = FALSE) +
  scale_color_manual(
    values = c("elasticity_TWFE" = "steelblue", "elasticity_new_TWFE" = "darkred"),
    labels = c("elasticity_TWFE" = "equal k", "elasticity_new_TWFE" = "heterogeneous k")
  ) +
  labs(
    x = "Log GDP per capita (PPP)",
    y = "Demand elasticity",
    color = NULL,
    title = "Income gradient in demand elasticity",
    subtitle = "Old vs loan-corrected elasticities"
  ) +
  theme_classic()
```

# Regression comparisons (TWFE vs OWFE vs Pooled)

Consider the following panel data models:

$$
Y_{it} = \beta_i D_{it} + \alpha_i + \varepsilon_{it} \quad \text{(One-way fixed effects)}
$$

$$
Y_{it} = \beta_i D_{it} + \alpha_i + \gamma_t + \varepsilon_{it} \quad \text{(Two-way fixed effects)}
$$

where $i$ indexes countries and $t$ indexes time. The parameter $\beta_i$ captures the country-specific treatment effect (elasticity), $\alpha_i$ are country fixed effects, and $\gamma_t$ are time fixed effects. The one-way fixed effects estimator removes country means:

$$
\tilde{Y}_{it}^{(1)} = Y_{it} - \bar{Y}_i
$$

The two-way fixed effects estimator removes both country and time means:

$$
\tilde{Y}_{it}^{(2)} = Y_{it} - \bar{Y}_i - \bar{Y}_t + \bar{Y}
$$

Subtracting the one-way transformation from the two-way transformation yields:

$$
\tilde{Y}_{it}^{(2)} - \tilde{Y}_{it}^{(1)} = -\bar{Y}_t + \bar{Y}
$$
$$
\Rightarrow \tilde{Y}_{it}^{(2)} = \tilde{Y}_{it}^{(1)} - (\bar{Y}_t - \bar{Y})
$$

If $\gamma_t$ represents a common time effect across countries, then:

$$
\bar{Y}_t - \bar{Y} \approx \gamma_t
$$

so that:

$$
\tilde{Y}_{it}^{(2)} = \tilde{Y}_{it}^{(1)} - \gamma_t
$$

Let $\hat{\beta}_i^{OWFE}$ and $\hat{\beta}_i^{TWFE}$ denote the country-specific elasticities obtained from the one-way and two-way fixed effects models, respectively. Then:

$$
\hat{\beta}_i^{TWFE} = \hat{\beta}_i^{OWFE} - c
$$

where $c$ is a constant given by the average time effect:

$$
c = \frac{1}{T} \sum_{t} \gamma_t
$$

This implies a linear relationship between the two estimators:

$$
\hat{\beta}_i^{TWFE} = \alpha + 1 \cdot \hat{\beta}_i^{OWFE}, \quad \text{with } \alpha = -c
$$

Thus, the slope is equal to one and the difference between the two estimators is a constant shift.

```{r regression-comparison}
# Convention throughout: "old" = equal k (common price change),
# "new" = heterogeneous k (country-specific loan-adjusted price change).
reg_old_TWFE <- lm(elasticity_TWFE          ~ log_gdp, data = elas_df)
reg_new_TWFE <- lm(elasticity_new_TWFE      ~ log_gdp, data = elas_df)
reg_old_controls <- lm(elasticity_controls      ~ log_gdp, data = elas_df)
reg_new_controls <- lm(elasticity_controls_new  ~ log_gdp, data = elas_df)

## NOTE: in your upload the pooled pair was labelled the other way round
## (reg_old_pooled used the heterogeneous-k column). Aligned here so old=equal k,
## new=heterogeneous k, matching the TWFE and OWFE pairs above. Both regressions
## are the same two you added; only the names were swapped to be consistent.
reg_old_pooled <- lm(elasticity_Pooled      ~ log_gdp, data = elas_df)
reg_new_pooled <- lm(elasticity_new_Pooled  ~ log_gdp, data = elas_df)

stargazer(reg_old_TWFE, reg_new_TWFE, reg_old_controls , reg_new_controls , type = "text")
stargazer(reg_old_pooled, reg_new_pooled, type = "text")
```

# Calibration robustness of the gradient collapse

The headline result is that the income gradient in the loan-corrected elasticity
(`reg_new_TWFE`: `elasticity_new_TWFE ~ log_gdp`) collapses to insignificance once
the effective price change is made heterogeneous. That result is computed at a single
calibration of the loan model, `r = 0.03` and `g = 0.02`. Because $k_i$ is a constructed
quantity rather than an estimated one, a referee will ask whether the collapse is real
or an artifact of where the loan parameters were set. Since $k_i$ falls with income by
construction, dividing by it will attenuate an income gradient almost mechanically, so
the burden is to show the collapse survives across defensible parameter choices and,
more importantly, that it is the loan structure doing the work rather than any
income-decreasing divisor. The next two sections address each in turn.

## Sweeping the loan parameters one at a time

For each value of the discount rate $r$ and the wage-growth rate $g$ across defensible
bands, we re-solve the present value, rebuild $k_i$ for every country, recompute the
loan-corrected TWFE elasticity, and re-run the income-gradient regression. Holding one
parameter at baseline while the other moves isolates the sensitivity to each.

```{r calib-sweep}
## Recompute the loan-corrected TWFE gradient for a given (r, g). Everything else
## (wages, Estimate_TWFE, the log_gdp used in the headline regression) is held fixed;
## only the loan parameters that feed k_i are varied.
gradient_at <- function(r_val, g_val) {
  d <- wages %>%
    mutate(
      PV_k  = vapply(wage_gbp, function(w)
        pv_loan(w, L = L, alpha = loan_alpha, tau = loan_tau,
                g = g_val, r = r_val, horizon = loan_T),
        numeric(1)),
      k_i_s = log(Ppost) - log(PV_k),
      eta_s = as.numeric(Estimate_TWFE) / k_i_s
    ) %>%
    left_join(elas_df %>% select(Country, log_gdp), by = "Country") %>%
    filter(!is.na(eta_s), !is.na(log_gdp), is.finite(eta_s))

  m  <- lm(eta_s ~ log_gdp, data = d)
  co <- summary(m)$coefficients["log_gdp", ]
  c(slope = unname(co["Estimate"]), p = unname(co["Pr(>|t|)"]), n = nrow(d))
}

r_base <- loan_r   # 0.03, baseline discount rate
g_base <- loan_g   # 0.02, baseline wage growth

# discount rate sweep, g held at baseline
r_grid  <- seq(0.01, 0.06, by = 0.005)
sweep_r <- as.data.frame(t(vapply(r_grid, function(rr) gradient_at(rr, g_base), numeric(3))))
sweep_r$param <- "Discount rate r"; sweep_r$value <- r_grid

# wage growth sweep, r held at baseline
g_grid  <- seq(0.00, 0.04, by = 0.005)
sweep_g <- as.data.frame(t(vapply(g_grid, function(gg) gradient_at(r_base, gg), numeric(3))))
sweep_g$param <- "Wage growth g"; sweep_g$value <- g_grid

sweep_all <- bind_rows(sweep_r, sweep_g)
names(sweep_all)[1:2] <- c("slope", "p")

# the table a referee scans: is the gradient ever significant across the sweep?
sweep_all %>%
  mutate(across(c(slope, p), ~ round(.x, 3)),
         significant_5pct = p < 0.05) %>%
  select(param, value, slope, p, significant_5pct, n)
```

```{r calib-sweep-plot, fig.width=7, fig.height=4}
ggplot(sweep_all, aes(value, p)) +
  geom_hline(yintercept = 0.05, linetype = "dashed", colour = "#8B0000", linewidth = 0.4) +
  geom_line(colour = accent, linewidth = 0.8) +
  geom_point(colour = accent, size = 2) +
  facet_wrap(~ param, scales = "free_x") +
  labs(
    x = "Parameter value",
    y = "p-value of the income gradient",
    title = "The gradient collapse is not an artifact of the loan parameters",
    subtitle = "Loan-corrected TWFE elasticity on log GDP per capita, one parameter swept at a time",
    caption = paste(
      "Notes. Each point re-solves the present value at that parameter, rebuilds k_i, recomputes the",
      "corrected elasticity, and re-runs the gradient regression. Dashed line is the 5% level; points",
      "above it mean the income gradient is not significant at that calibration.",
      sep = "\n"
    )
  ) +
  theme_eer()
```

Read this honestly. If the p-value stays above 0.05 across the whole sweep, the collapse
is robust to the loan calibration and the "arbitrary parameter" objection is closed. If
the gradient becomes significant in part of the range, that region has to be reported and
the headline claim qualified to the parameter values where it holds, rather than stated
unconditionally.

## Placebo denominator: is it the loan structure or just income?

The sweep shows robustness to the parameters, but not the deeper question. Because $k_i$
decreases with income, *any* income-decreasing divisor would flatten an income gradient.
The test that actually adjudicates the mechanism is to divide the same $\beta_i$ by a
placebo denominator that is a pure function of income and carries none of the loan
structure, then ask whether the gradient still collapses. If it collapses under the
placebo too, the loan story adds nothing and the critics are right that the result is
mechanical. If it collapses under the real $k_i$ but *not* under the placebo, the loan
valuation is doing work a naive income deflator cannot replicate.

```{r placebo-denominator}
## Build a placebo k that is a monotone function of income only, scaled to sit on the
## same range as the real k_i so the comparison is like-for-like. It has no threshold,
## no repayment rule, no discounting: it is income and nothing else.
placebo_df <- elas_df %>%
  filter(!is.na(k_i), !is.na(Estimate_TWFE), !is.na(log_gdp)) %>%
  mutate(
    # a pure log-income divisor, rescaled to the observed [min, max] of the real k_i
    k_income_raw = log(wage_gbp),
    k_placebo    = (k_income_raw - min(k_income_raw)) /
                   (max(k_income_raw) - min(k_income_raw)) *
                   (max(k_i) - min(k_i)) + min(k_i),
    eta_placebo  = as.numeric(Estimate_TWFE) / k_placebo
  )

# three gradients side by side: equal k, real heterogeneous k, placebo income-only k
g_equal   <- lm(elasticity_TWFE     ~ log_gdp, data = placebo_df)
g_real    <- lm(elasticity_new_TWFE ~ log_gdp, data = placebo_df)
g_placebo <- lm(eta_placebo         ~ log_gdp, data = placebo_df)

placebo_summary <- bind_rows(
  tidy(g_equal,   conf.int = TRUE) %>% filter(term == "log_gdp") %>% mutate(denominator = "Equal k (common price change)"),
  tidy(g_real,    conf.int = TRUE) %>% filter(term == "log_gdp") %>% mutate(denominator = "Real k_i (loan-adjusted)"),
  tidy(g_placebo, conf.int = TRUE) %>% filter(term == "log_gdp") %>% mutate(denominator = "Placebo k (income only)")
) %>%
  transmute(denominator,
            slope    = round(estimate, 3),
            conf.low = round(conf.low, 3),
            conf.high= round(conf.high, 3),
            p        = round(p.value, 3),
            significant_5pct = p.value < 0.05)

placebo_summary
```

```{r placebo-plot, fig.width=7, fig.height=3.5}
placebo_summary %>%
  mutate(denominator = factor(denominator,
           levels = c("Equal k (common price change)",
                      "Real k_i (loan-adjusted)",
                      "Placebo k (income only)"))) %>%
  ggplot(aes(slope, denominator)) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey60", linewidth = 0.4) +
  geom_pointrange(aes(xmin = conf.low, xmax = conf.high, colour = significant_5pct),
                  linewidth = 0.8, size = 0.6) +
  scale_colour_manual(values = c(`TRUE` = "#8B0000", `FALSE` = "grey55"), guide = "none") +
  labs(
    x = "Income gradient (slope on log GDP per capita)",
    y = NULL,
    title = "Is it the loan, or just income? The placebo denominator answers",
    subtitle = "Gradient of the elasticity under three denominators, with 95% intervals",
    caption = paste(
      "Notes. Equal k divides by the common log price change. Real k_i divides by the loan-adjusted",
      "price change. Placebo k divides by a pure income function rescaled to the same range as k_i.",
      "If the gradient collapses under real k_i but not under placebo k, the loan structure is doing",
      "the work rather than income alone. Crimson marks significance at the 5% level.",
      sep = "\n"
    )
  ) +
  theme_eer()
```

The decisive comparison is the bottom two rows. The equal-$k$ gradient is the uncorrected
benchmark and is expected to be steep and significant. If the real-$k_i$ gradient is flat
and insignificant while the placebo-$k$ gradient remains steep and significant, the paper
can claim that the loan valuation, and not a generic income rescaling, is what dissolves
the gradient. If instead the placebo also flattens the gradient, that is evidence the
result is mechanical, and it should be reported as such rather than suppressed.

# Other interesting correlations

```{r other-correlations}
# Reuse the single WB frame already pulled above (elas_df carries the basket).
wb_vars <- c("gdp_ppp_pc", "gini", "youth_unemp",
             "educ_spend_gdp", "tertiary_enroll", "net_migration")

# elas_df renamed gdp_ppp_pc to GDP_PPP, so map that back for the GDP column
dat <- elas_df %>% rename(gdp_ppp_pc = GDP_PPP)

min_n <- 8   # minimum complete cases to include a model

models_new <- list()
kept_vars  <- c()

for (v in wb_vars) {
  dsub <- dat %>%
    select(elasticity_new_TWFE, all_of(v)) %>%
    filter(complete.cases(.))

  if (nrow(dsub) < min_n) next                  # too few observations
  if (sd(dsub[[v]], na.rm = TRUE) == 0) next     # no variation in regressor

  models_new[[v]] <- lm(reformulate(v, response = "elasticity_new_TWFE"),
                        data = dsub)
  kept_vars <- c(kept_vars, v)
}

print(kept_vars)
stargazer(models_new, type = "text")
```
```{r}
library(broom)

# readable labels for the six indicators
nice_names <- c(
  gdp_ppp_pc      = "GDP per capita (PPP)",
  gini            = "Gini",
  youth_unemp     = "Youth unemployment",
  educ_spend_gdp  = "Education spending (% GDP)",
  tertiary_enroll = "Tertiary enrollment",
  net_migration   = "Net migration"
)

wb_vars <- names(nice_names)

# refit each bivariate model on standardized variables so slopes are comparable,
# and carry the R-squared along
std_df <- bind_rows(lapply(wb_vars, function(v) {
  d <- dat %>% select(y = elasticity_new_TWFE, x = all_of(v)) %>% filter(complete.cases(.))
  if (nrow(d) < 8 || sd(d$x) == 0) return(NULL)
  d <- d %>% mutate(x = as.numeric(scale(x)), y = as.numeric(scale(y)))
  m <- lm(y ~ x, data = d)
  tidy(m, conf.int = TRUE) %>%
    filter(term == "x") %>%
    mutate(variable = nice_names[v], r2 = summary(m)$r.squared, n = nrow(d))
})) %>%
  mutate(
    sig = p.value < 0.05,
    variable = factor(variable, levels = variable[order(estimate)])
  )

theme_eer <- function(base_size = 12) {
  theme_classic(base_size = base_size) +
    theme(
      text             = element_text(colour = "grey15"),
      plot.title       = element_text(size = base_size + 1, face = "plain"),
      plot.subtitle    = element_text(size = base_size - 1, colour = "grey35"),
      plot.caption     = element_text(size = base_size - 3, colour = "grey40",
                                      hjust = 0, margin = margin(t = 8)),
      axis.line          = element_line(colour = "grey20", linewidth = 0.4),
      axis.ticks         = element_line(colour = "grey20", linewidth = 0.4),
      panel.grid.major.y = element_line(colour = "grey92", linewidth = 0.3),
      legend.position    = "bottom",
      legend.key.width   = unit(1.4, "lines")
    )
}

accent <- "#6A0DAD"
ggplot(std_df, aes(estimate, variable)) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey60", linewidth = 0.4) +
  geom_pointrange(aes(xmin = conf.low, xmax = conf.high, colour = sig),
                  linewidth = 0.8, size = 0.6) +
  geom_text(aes(label = sprintf("R\u00B2 = %.2f", r2)),
            hjust = -0.25, vjust = -1, size = 3, colour = "grey35") +
  scale_colour_manual(values = c(`TRUE` = "#8B0000", `FALSE` = "grey65"),
                      guide = "none") +
  scale_x_continuous(expand = expansion(mult = c(0.08, 0.18))) +
  labs(
    x = "Standardized association with loan-corrected elasticity",
    y = NULL,
    title = "Only education spending moves the loan-corrected elasticity",
    subtitle = "Standardized bivariate slopes with 95% intervals, one regressor at a time",
    caption = paste(
      "Notes. Each row is a separate OLS regression of the loan-corrected TWFE elasticity",
      "on one standardized indicator. Crimson marks significance at the 5% level.",
      sep = "\n"
    )
  ) +
  theme_eer()
```

# Publication figures

A small set of journal-style figures in the restrained black-and-grey register common to the empirical economics of education literature. One shared theme, one accent colour, light horizontal guides only, and a notes line under each figure rather than a heavy title.

```{r eer-theme}
# one shared theme for every figure below
theme_eer <- function(base_size = 12) {
  theme_classic(base_size = base_size) +
    theme(
      text             = element_text(colour = "grey15"),
      plot.title       = element_text(size = base_size + 1, face = "plain"),
      plot.subtitle    = element_text(size = base_size - 1, colour = "grey35"),
      plot.caption     = element_text(size = base_size - 3, colour = "grey40",
                                      hjust = 0, margin = margin(t = 8)),
      axis.line        = element_line(colour = "grey20", linewidth = 0.4),
      axis.ticks       = element_line(colour = "grey20", linewidth = 0.4),
      panel.grid.major.y = element_line(colour = "grey92", linewidth = 0.3),
      legend.position  = "bottom",
      legend.key.width = unit(1.4, "lines")
    )
}

accent <- "#1F4E79"   # one muted navy accent, prints cleanly in greyscale
```

## Figure 1. The loan correction shifts each country's elasticity

A dumbbell figure. For every country the open grey point is the elasticity under a single common price change, and the filled navy point is the elasticity once the price change is adjusted for the country-specific present value of the loan. The segment is the size of the correction.

```{r fig-dumbbell, fig.width=7, fig.height=6}
dumbbell_df <- elas_df %>%
  transmute(Country,
            equal_k  = elasticity_TWFE,
            hetero_k = elasticity_new_TWFE) %>%
  filter(!is.na(equal_k), !is.na(hetero_k)) %>%
  arrange(hetero_k) %>%
  mutate(Country = factor(Country, levels = Country))

ggplot(dumbbell_df) +
  geom_segment(aes(x = equal_k, xend = hetero_k, y = Country, yend = Country),
               colour = "grey70", linewidth = 0.6) +
  geom_point(aes(equal_k,  Country, shape = "Equal k"),
             colour = "grey55", size = 2.6, fill = "white", stroke = 0.7) +
  geom_point(aes(hetero_k, Country, shape = "Heterogeneous k"),
             colour = accent, size = 2.6) +
  scale_shape_manual(values = c("Equal k" = 21, "Heterogeneous k" = 16), name = NULL) +
  labs(
    x = "Implied demand elasticity (TWFE)",
    y = NULL,
    title = "Loan adjustment changes the implied elasticity country by country",
    caption = paste(
      "Notes. Equal k applies a single log price change, log(22200) minus log(9250).",
      "Heterogeneous k applies each country's loan-adjusted price change k_i.",
      "Countries are ordered by the heterogeneous-k elasticity."
    )
  ) +
  theme_eer()
```

## Figure 2. Income gradient in the loan-corrected elasticity

The loan-corrected elasticity against log GDP per capita, with country labels, an OLS fit, and a confidence band. This is the headline correlation the regression table formalises.

```{r fig-gradient}
shift_df <- elas_df %>%
  select(Country, log_gdp, old = elasticity_TWFE, new = elasticity_new_TWFE) %>%
  filter(!is.na(old), !is.na(new), !is.na(log_gdp)) %>%
  arrange(log_gdp) %>%
  mutate(x = row_number())          # one distinct column per country, ordered by income

ggplot(shift_df, aes(x = x)) +

 # each country's shift, arrow running from the old to the corrected elasticity,
  # coloured by income so the columns carry the GDP dimension
  geom_segment(aes(xend = x, y = old, yend = new, colour = log_gdp),
               arrow = arrow(length = unit(0.06, "inches"), type = "closed"),
               linewidth = 0.8) +

  geom_point(aes(y = old), shape = 21, fill = "white", colour = "grey55", size = 2) +
  geom_point(aes(y = new, colour = log_gdp), size = 2.4) +

  # the two gradients laid over the ordered columns
  geom_smooth(aes(y = old), method = "lm", se = FALSE,
              colour = "grey55", linetype = "dashed", linewidth = 0.8) +
  geom_smooth(aes(y = new), method = "lm", se = FALSE,
              colour = "#6A0DAD", linewidth = 0.9) +

  scale_colour_gradient(low = "#8B0000", high = "#4B0082",
                        name = "Log GDP per capita") +
  scale_x_continuous(breaks = shift_df$x, labels = shift_df$Country,
                     expand = expansion(add = 0.6)) +

  labs(
    x = NULL,
    y = "Demand elasticity (TWFE)",
    title = "Each country's correction, and the gradient it reverses",
    subtitle = "Countries ordered by income, arrows run from equal-k to loan-corrected elasticity",
caption = paste(
      "Notes. Dashed grey is the OLS gradient of the uncorrected elasticity over the income",
      "ordering, which slopes up. Solid navy is the gradient of the loan-corrected elasticity,",
      "which is flat. Arrows run from equal k to heterogeneous k, coloured by income.",
      sep = "\n"
    )
    ) +
  theme_eer() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8))
```

## Figure 3. Income-gradient slope across specifications

A coefficient figure. Each row is the slope on log GDP per capita from one elasticity specification, with a 95% interval. It shows whether the gradient survives the move from equal to heterogeneous k and across the fixed-effects designs.

```{r fig-coef, fig.width=7, fig.height=4}
slope_df <- bind_rows(
  tidy(reg_old_TWFE,   conf.int = TRUE) %>% mutate(model = "TWFE, equal k"),
  tidy(reg_new_TWFE,   conf.int = TRUE) %>% mutate(model = "TWFE, heterogeneous k"),
  tidy(reg_old_controls,   conf.int = TRUE) %>% mutate(model = "TWFE + controls, equal k"),
  tidy(reg_new_controls,   conf.int = TRUE) %>% mutate(model = "TWFE + controls, heterogeneous k"),
  tidy(reg_old_pooled, conf.int = TRUE) %>% mutate(model = "Pooled, equal k"),
  tidy(reg_new_pooled, conf.int = TRUE) %>% mutate(model = "Pooled, heterogeneous k")
) %>%
  filter(term == "log_gdp") %>%
  mutate(model = factor(model, levels = rev(model)))

ggplot(slope_df, aes(estimate, model)) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey60", linewidth = 0.4) +
  geom_pointrange(aes(xmin = conf.low, xmax = conf.high),
                  colour = accent, linewidth = 0.6, size = 0.5) +
  labs(
    x = "Slope on log GDP per capita",
    y = NULL,
    title = "The income gradient holds across specifications",
    caption = "Notes. Points are OLS slope coefficients with 95% confidence intervals."
  ) +
  theme_eer()
```

## Figure 4. How much the loan compresses the effective price increase

The naive reform raises the log price by k_common for everyone. The loan shrinks that for countries whose graduates expect to repay little. This ranks each country by its loan-adjusted price change k_i against the common benchmark.

```{r fig-kbar, fig.width=7, fig.height=6}
k_df <- elas_df %>%
  select(Country, k_i) %>%
  filter(!is.na(k_i)) %>%
  arrange(k_i) %>%
  mutate(Country = factor(Country, levels = Country))

ggplot(k_df, aes(k_i, Country)) +
  geom_segment(aes(x = k_common, xend = k_i, y = Country, yend = Country),
               colour = "grey75", linewidth = 0.5) +
  geom_vline(xintercept = k_common, linetype = "dashed",
             colour = "grey45", linewidth = 0.5) +
  geom_point(colour = accent, size = 2.4) +
  annotate("text", x = k_common, y = Inf, label = "Common k",
           vjust = 1.6, hjust = -0.05, size = 3, colour = "grey45") +
  labs(
    x = "Loan-adjusted log price change, k_i",
    y = NULL,
    title = "The loan compresses the effective price increase, unevenly",
    caption = "Notes. Dashed line is the common log price change, log(22200) minus log(9250). Each point is a country's loan-adjusted k_i."
  ) +
  theme_eer()
```
