---
title: "Synthetic controls — weighted construction of the counterfactual control group"
author: "annon for peer review"
date: "2025-11-22"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# What this script does

This script turns the methodology in the accompanying note into runnable R.
It moves through five stages and each stage has a short explanation, the matching
equation, and then the code that implements it.

The five stages are pre-treatment trajectories, within-country standardisation
and distance, donor screening (the donor pool), constrained synthetic-control
weights, and finally the global donor weights. The last chunk writes a single
Excel workbook with two sheets, one holding the donor pool and one holding the
weights.

```{r packages}
# install once if needed:
# install.packages(c("readxl","dplyr","tidyr","writexl","Rsolnp"))

library(readxl)   # read the source spreadsheet
library(dplyr)    # data wrangling
library(tidyr)    # reshape long <-> wide
library(writexl)  # write the final workbook
library(Rsolnp)   # constrained optimiser for the convex weights
```

# Data and the pre-treatment window

We read the filtered UCAS data and keep only the pre-Brexit years, meaning every
observation from 2020 or earlier. All donor screening and all weight estimation
happen inside this window, so nothing post-treatment can leak into the
construction of the control group.

The pre-treatment trajectory of a country $c$ is just its sequence of
application counts up to 2020.

$$
Y_c^{\text{pre}} = \bigl(\text{App}_{c,2014},\ \text{App}_{c,2015},\ \ldots,\ \text{App}_{c,2020}\bigr)
$$

```{r read-data}
data <- read_excel("data_filtered.xlsx", sheet = "data_filtered")

# the 26 treated EU origin markets
eu_real <- c(
  "Austria", "Belgium", "Bulgaria", "Croatia", "Cyprus (European Union)",
  "Czech Republic", "Denmark", "Estonia", "Finland", "France",
  "Germany", "Greece", "Hungary", "Italy",
  "Latvia", "Lithuania", "Luxembourg", "Malta", "Netherlands",
  "Poland", "Portugal", "Romania", "Slovakia", "Slovenia",
  "Spain", "Sweden"
)

# keep only the pre-treatment window
data <- data %>% filter(Year <= 2020)

# how many donors each EU market is matched to
n_donors <- 5
```

# Building the country-by-year trajectory matrix

We collapse the data to one row per country and one column per year. Each cell is
the total number of applicants for that country in that year. This wide matrix is
the raw material for every distance and every fit that follows.

```{r build-matrix}
pre <- data %>%
  select(Domicile_named_country, Year, Applicants)

traj <- pre %>%
  group_by(Domicile_named_country, Year) %>%
  summarise(Applicants = sum(Applicants), .groups = "drop") %>%
  pivot_wider(names_from = Year, values_from = Applicants) %>%
  as.data.frame()

rownames(traj) <- traj$Domicile_named_country
traj <- traj[, -1]            # drop the name column, keep years only

# logged levels, reused later for the actual fit
logtraj <- log(traj + 1)
```

# Within-country standardisation

Donor screening should depend on the shape of a country's path, not on its size.
A large market and a small market that grow in the same pattern should look
similar. To get that, we log each series and then standardise it inside each
country, subtracting that country's own mean and dividing by its own standard
deviation over the pre-period.

$$
Z_{c,t} = \frac{\log(\text{App}_{c,t} + 1) - \mu_c}{\sigma_c}
$$

Here $\mu_c$ and $\sigma_c$ are computed across the years of country $c$, not
across countries. This is the one place the original script needed a fix. The
base `scale()` function standardises columns, which would have standardised each
year across countries. We want the opposite, so we transpose, scale, and
transpose back, which standardises each country's own row.

```{r standardise}
# t(scale(t(.))) standardises ROW-WISE, i.e. within each country
traj_scaled <- t(scale(t(logtraj)))
```

# Pairwise distance and the donor pool

With every country reduced to a standardised shape, we measure how close two
shapes are with Euclidean distance over the pre-period.

$$
d_{j,k} = \sqrt{\sum_{t \le 2020}\bigl(Z_{j,t} - Z_{k,t}\bigr)^2}
$$

For each treated EU market $j$ we then keep the five non-EU origins with the
smallest distance. That set is the donor pool $\Gamma_j$.

$$
\Gamma_j = \operatorname*{arg\,min}_{k \notin \text{EU}}{}^{(5)} \ d_{j,k},
\qquad |\Gamma_j| = 5
$$

```{r distances}
dist_mat <- as.matrix(dist(traj_scaled, method = "euclidean"))

treated <- rownames(traj_scaled) %in% eu_real
donors  <- !treated

# rows = treated EU markets, columns = candidate non-EU donors
dist_treated_donors <- dist_mat[treated, donors, drop = FALSE]
eu_present <- rownames(dist_treated_donors)
```

# Constrained synthetic-control weights

Screening tells us which donors are eligible. It does not tell us how to combine
them. For that we leave the standardised shapes behind and go back to logged
levels, then find the convex combination of the five donors that best tracks the
EU market over the pre-period.

$$
w_j = \operatorname*{arg\,min}_{w_j}\
\sum_{t \le 2020}\Bigl(\log(\text{App}_{j,t}+1) - \sum_{k \in \Gamma_j} w_{jk}\,\log(\text{App}_{k,t}+1)\Bigr)^2
$$

The combination is restricted to be a genuine weighted average, so weights are
non-negative and add up to one. Non-negativity stops the fit from extrapolating
beyond the donors we actually observe, and the adding-up rule keeps the synthetic
market inside the range of real donor paths.

$$
w_{jk} \ge 0, \qquad \sum_{k \in \Gamma_j} w_{jk} = 1
$$

The helper below states the problem in exactly those terms. The objective is the
sum of squared residuals, the equality constraint is the sum of weights, and the
bounds give non-negativity. `Rsolnp::solnp` reads almost like the equation.

```{r weight-helper}
solve_weights <- function(y, Z) {
  n <- ncol(Z)
  fit <- solnp(
    pars  = rep(1 / n, n),                       # start at equal weights
    fun   = function(w) sum((y - Z %*% w)^2),    # squared prediction error
    eqfun = function(w) sum(w),                  # adding-up ...
    eqB   = 1,                                   # ... must equal one
    LB    = rep(0, n),                           # non-negativity
    UB    = rep(1, n),
    control = list(trace = 0)
  )
  as.numeric(fit$pars)
}
```

We loop over the EU markets. For each one we take its five donors, restrict to the
years where the market and all five donors are observed, solve for the weights,
build the fitted path, and record the root mean squared prediction error.

$$
\widehat{Y}_{j,t} = \sum_{k \in \Gamma_j} w_{jk}\,\log(\text{App}_{k,t}+1),
\qquad
\text{RMSPE}_j = \sqrt{\frac{1}{T_0}\sum_{t \le 2020}\bigl(\log(\text{App}_{j,t}+1) - \widehat{Y}_{j,t}\bigr)^2}
$$

```{r fit-loop}
logmat <- as.matrix(logtraj)   # rows = countries, cols = years
rows <- list()                 # collect one row per (EU market, donor) pair

for (j in eu_present) {

  # five nearest donors for this EU market
  d_row       <- dist_treated_donors[j, ]
  donor_names <- names(sort(d_row))[1:n_donors]

  # years where the market and all donors are observed
  block <- logmat[c(j, donor_names), , drop = FALSE]
  good  <- colSums(is.na(block)) == 0
  yrs   <- colnames(block)[good]

  y <- as.numeric(logmat[j, yrs])
  Z <- t(logmat[donor_names, yrs, drop = FALSE])   # rows = years, cols = donors

  w   <- solve_weights(y, Z)
  fit <- as.numeric(Z %*% w)
  rmspe <- sqrt(mean((y - fit)^2))

  rows[[j]] <- data.frame(
    EU_country   = j,
    donor_rank   = seq_len(n_donors),
    donor        = donor_names,
    distance     = as.numeric(d_row[donor_names]),
    synth_weight = w,
    rmspe        = rmspe,
    stringsAsFactors = FALSE
  )
}

donor_pool <- do.call(rbind, rows)
rownames(donor_pool) <- NULL
head(donor_pool, n_donors)
```

# Aggregating to one global weight per donor

The difference-in-differences model uses a single common control path for every
EU market, so the country-specific weights have to be collapsed into one weight
per donor. We average each donor's weight across the treated markets, treating a
donor that never enters a pool as a zero for that market.

$$
w_k = \frac{1}{J}\sum_{j=1}^{J} w_{jk}
$$

We then normalise so the global weights sum to one. Because each EU market already
contributed weights that summed to one, this normalisation is essentially a safety
step and barely moves the numbers.

$$
\tilde{w}_k = \frac{w_k}{\sum_s w_s}
$$

Finally, to give the control side the same total influence as the treated side in
the weighted regression, the normalised donor weights are scaled by the number of
treated markets $J$, while every EU observation carries unit weight.

$$
\omega_i =
\begin{cases}
1 & i \in \Omega \ \text{(treated EU)} \\
J \cdot \tilde{w}_k & i \in P \ \text{(donor)}
\end{cases}
$$

```{r global-weights}
J <- length(eu_present)   # number of treated markets that got a synthetic control

weights <- donor_pool %>%
  group_by(donor) %>%
  summarise(w_k = sum(synth_weight) / J, .groups = "drop") %>%
  mutate(
    normalized_weight  = w_k / sum(w_k),     # w-tilde
    regression_weight  = J * normalized_weight
  ) %>%
  rename(donor_country = donor, global_weight = w_k) %>%
  arrange(desc(normalized_weight))

weights
```

# Writing the two-sheet workbook

The donor pool and the weights go into one Excel file with two sheets. The first
sheet lists, for every EU market, its five donors with their distance, the convex
weight each donor receives, and the market's pre-treatment RMSPE. The second sheet
gives the one global weight per donor along with the normalised weight and the
regression weight used in the weighted model.

```{r write-excel}
write_xlsx(
  list(
    donor_pool = donor_pool,
    weights    = weights
  ),
  path = "control_group_output.xlsx"
)
```

# Identification safeguards

A short checklist mirrors the safeguards in the methodology note. Donor selection
and weight estimation use only observations from 2020 or earlier. The estimated
weights stay constant across the full panel once the model is run. Pre-treatment
fit should look acceptable in both trajectory plots and the RMSPE column produced
above. The weighted specification should be reported next to the unweighted
baseline so readers can judge sensitivity to how the control group was built. And
because the reported standard errors condition on the estimated weights, a
bootstrap that rebuilds the donor pools and the weights in every replication makes
a sensible extra inference check.
