---
title: "Synthetic controls — weighted construction of the counterfactual control group"
author: "annon for peer review"
date: "2025-11-22"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# What this script does

This script turns the methodology in the accompanying note into runnable
R. It moves through five stages and each stage has a short explanation,
the matching equation, and then the code that implements it.

The five stages are pre-treatment trajectories, within-country
standardisation and distance, donor screening (the donor pool),
constrained synthetic-control weights, and finally the global donor
weights. The last chunk writes a single Excel workbook with two sheets,
one holding the donor pool and one holding the weights.

```{r packages}


library(readxl)   
library(dplyr)   
library(tidyr)    
library(writexl) 
library(Rsolnp)   
```

# Data and the pre-treatment window

We read the filtered UCAS data and keep the seven pre-Brexit years from
2014 through 2020. All donor screening and
all weight estimation happen inside this window, so nothing
post-treatment can leak into the construction of the control group.

The pre-treatment trajectory of a country $c$ is just its sequence of
application counts up to 2020.

$$
Y_c^{\text{pre}} = \bigl(\text{App}_{c,2014},\ \text{App}_{c,2015},\ \ldots,\ \text{App}_{c,2020}\bigr)
$$

```{r read-data}
data <- read_excel("data_filtered_FOR_FIRST_ONLY.xlsx", sheet = "full_data")

required_columns <- c("Domicile_named_country", "Year", "Applicants")
stopifnot(all(required_columns %in% names(data)))

# entities that are NOT sovereign foreign countries
non_countries <- c(
  # UK home nations
  "England", "Scotland", "Wales", "Northern Ireland", 
  # Crown Dependencies
  "Alderney", "Guernsey", "Jersey", "Isle of Man",
  # British Overseas Territories
  "Anguilla", "Bermuda", "British Antarctic Territory",
  "British Indian Ocean Territory", "British Virgin Islands",
  "Cayman Islands", "Falkland Islands (Malvinas)", "Gibraltar",
  "Montserrat", "St Helena, Ascension & Tristan da Cunha",
  "Turks and Caicos Islands",
  # dependencies / overseas regions of other states
  "Aland Islands", "Aruba", "Canary Islands", "Christmas Island",
  "Curacao", "Faroe Islands", "French Guiana", "French Polynesia",
  "Greenland", "Guadeloupe", "Guam", "Hong Kong", "Macao",
  "Martinique", "Netherlands Antilles", "New Caledonia",
  "Norfolk Island", "Northern Mariana Islands", "Puerto Rico",
  "Reunion", "Sint Maarten (Dutch Part)", "St Barthelemy",
  "St Martin (French part)", "Virgin Islands (US)",
  # non-standard Cyprus codings (Cyprus is already treated via eu_real)
  "Cyprus (Non-European Union)", "Cyprus (Not otherwise specified)",
  # aggregates, placeholders, defunct codes
  "All", "No information provided (Overseas)", "Not Known",
  "Stateless", "USSR (Not In Use)", "French West Indies (Not In Use)"
)

data <- data %>%
  filter(!Domicile_named_country %in% non_countries)

# After removing the duplicated aggregate rows, the input must be a genuine
# country-year panel. Summing below is then a defensive aggregation, not a way
# to conceal duplicate records.
stopifnot(!anyDuplicated(data[c("Domicile_named_country", "Year")]))

# the 26 treated EU origin markets
eu_real <- c(
  "Austria", "Belgium", "Bulgaria", "Croatia", "Cyprus (European Union)",
  "Czech Republic", "Denmark", "Estonia", "Finland", "France",
  "Germany", "Greece", "Hungary", "Italy",
  "Latvia", "Lithuania", "Luxembourg", "Malta", "Netherlands",
  "Poland", "Portugal", "Romania", "Slovakia", "Slovenia",
  "Spain", "Sweden"
)

# keep only the seven-year pre-treatment window
expected_years <- 2014:2020
data <- data %>% filter(Year %in% expected_years)

data <- data %>%
  group_by(Domicile_named_country) %>%
  filter(
    n_distinct(Year) == length(expected_years),
    all(expected_years %in% Year),
    all(is.finite(Applicants)),
    all(Applicants >= 100)
  ) %>%
  ungroup()

```

# Building the country-by-year trajectory matrix

We collapse the data to one row per country and one column per year.
Each cell is the total number of applicants for that country in that
year. This wide matrix is the raw material for every distance and every
fit that follows.

```{r build-matrix}
pre <- data %>%
  select(Domicile_named_country, Year, Applicants)

traj <- pre %>%
  group_by(Domicile_named_country, Year) %>%
  summarise(Applicants = sum(Applicants), .groups = "drop") %>%
  pivot_wider(names_from = Year, values_from = Applicants) %>%
  as.data.frame()

rownames(traj) <- traj$Domicile_named_country
traj <- traj[, -1]            # drop the name column, keep years only
traj <- traj[, as.character(expected_years), drop = FALSE]

# Logged levels, reused later for donor screening and the synthetic fit.
# Use the ordinary logarithm of the observed applicant count.
logmat <- log(as.matrix(traj))

# log(0) is -Inf and log(negative values) is NaN. Keep only countries
# with finite logged values in every pre-treatment year.
valid_log_rows <- apply(logmat, 1, function(x) all(is.finite(x)))
logmat <- logmat[valid_log_rows, , drop = FALSE]

# A constant pre-treatment trajectory has zero within-country standard
# deviation and cannot be standardised for distance calculations.
row_log_sd <- apply(logmat, 1, sd)
valid_scale_rows <- is.finite(row_log_sd) & row_log_sd > 0
logmat <- logmat[valid_scale_rows, , drop = FALSE]

if (nrow(logmat) == 0L) {
  stop("No countries have complete, positive, non-constant pre-treatment trajectories.")
}

stopifnot(
  identical(colnames(logmat), as.character(expected_years)),
  all(eu_real %in% rownames(logmat))
)

# The standardisation section historically uses the name logtraj.
# Keep it as an explicit alias so both downstream object names exist.
logtraj <- logmat
```

```{r}
#explore the dimensions of the final database 

dim(logtraj)
```

The resulting sample contains `r nrow(logmat)` countries observed over
`r ncol(logmat)` pre-treatment years

# Within-country standardisation

Donor screening should depend on the shape of a country's path, not on
its size. A large market and a small market that grow in the same
pattern should look similar. To get that, we log each series and then
standardise it inside each country, subtracting that country's own mean
and dividing by its own standard deviation over the pre-period.

$$
Z_{c,t} = \frac{\log(\text{App}_{c,t}) - \mu_c}{\sigma_c}
$$

Here $\mu_c$ and $\sigma_c$ are computed across the years of country
$c$, not across countries. This is the one place the original script
needed a fix. The base `scale()` function standardises columns, which
would have standardised each year across countries. We want the
opposite, so we transpose, scale, and transpose back, which standardises
each country's own row. The code uses R's sample standard deviation,
whose denominator is $T_0-1$.

```{r standardise}
# t(scale(t(.))) standardises ROW-WISE, i.e. within each country
traj_scaled <- t(scale(t(logtraj)))

```

```{r}
# Convert matrices to data frames and retain row names as a column
#logtraj_df <- as.data.frame(logtraj)
#logtraj_df$country <- rownames(logtraj_df)
#logtraj_df <- logtraj_df[, c("country", setdiff(names(logtraj_df), "country"))]

#traj_scaled_df <- as.data.frame(traj_scaled)
#traj_scaled_df$country <- rownames(traj_scaled_df)
#traj_scaled_df <- traj_scaled_df[, c("country", setdiff(names(traj_scaled_df), "country"))]

# Export to Excel to test for correctness in subsample of countries (DONE)
#write_xlsx(
#  list(
#    logtraj = logtraj_df,
#    traj_scaled = traj_scaled_df
#  ),
#  path = "log_tests.xlsx"
#)
```

# Pairwise distance and the donor pool

With every country reduced to a standardised shape, we measure how close
two shapes are with Euclidean distance over the pre-period.

$$
d_{j,k} = \sqrt{\sum_{t \le 2020}\bigl(Z_{j,t} - Z_{k,t}\bigr)^2}
$$

For each treated EU market $j$ we then keep the three non-EU origins with
the smallest distance. That set is the donor pool $\Gamma_j$.

$$
\Gamma_j = \operatorname*{arg\,min}_{k \notin \text{EU}}{}^{(3)} \ d_{j,k},
\qquad |\Gamma_j| = 3
$$

```{r distances}
# Number of nearest non-EU donors retained for each treated EU market.
n_donors <- 3L

dist_mat <- as.matrix(dist(traj_scaled, method = "euclidean")) #change to euclidean otherwise

treated <- rownames(traj_scaled) %in% eu_real
donors  <- !treated

# rows = treated EU markets, columns = candidate non-EU donors
dist_treated_donors <- dist_mat[treated, donors, drop = FALSE]
eu_present <- rownames(dist_treated_donors)

if (length(eu_present) == 0L) {
  stop("No treated EU countries remain after constructing the logged panel.")
}

if (ncol(dist_treated_donors) < n_donors) {
  stop(sprintf(
    "Only %s eligible donors remain, fewer than n_donors = %s.",
    ncol(dist_treated_donors), n_donors
  ))
}
```

```{r}

# to manually verify everything we run this

#mat_to_df <- function(x, rowname_col = "country") {
#  data.frame(
#    country = rownames(x),
#    as.data.frame(x),
#    row.names = NULL,
#    check.names = FALSE
#  )
#}

#write_xlsx(
 # list(
#    traj_scaled = mat_to_df(traj_scaled),
#    dist_mat = mat_to_df(dist_mat)
# ),
 # path = "matrix_distances.xlsx"
#)
```

# Constrained synthetic-control weights

Screening tells us which donors are eligible. It does not tell us how to
combine them. For that we leave the standardised shapes behind and go
back to logged levels, then find the convex combination of the three
donors that best tracks the EU market over the pre-period.

$$
w_j = \operatorname*{arg\,min}_{w_j}\
\sum_{t \le 2020}\Bigl(\log(\text{App}_{j,t}) - \sum_{k \in \Gamma_j} w_{jk}\,\log(\text{App}_{k,t})\Bigr)^2
$$

The combination is restricted to be a genuine weighted average, so
weights are non-negative and add up to one. Non-negativity stops the fit
from extrapolating beyond the donors we actually observe, and the
adding-up rule keeps the synthetic market inside the range of real donor
paths.

$$
w_{jk} \ge 0, \qquad \sum_{k \in \Gamma_j} w_{jk} = 1
$$

The helper below states the problem in exactly those terms. The
objective is the sum of squared residuals, the equality constraint is
the sum of weights, and the bounds give non-negativity. `Rsolnp::solnp`
reads almost like the equation.

```{r weight-helper}
solve_weights <- function(y, Z) {
  n <- ncol(Z)
  fit <- solnp(
    pars  = rep(1 / n, n),                       # start at equal weights
    fun   = function(w) sum((y - Z %*% w)^2),    # squared prediction error
    eqfun = function(w) sum(w),                  # adding-up ...
    eqB   = 1,                                   # ... must equal one
    LB    = rep(0, n),                           # non-negativity
    UB    = rep(1, n),
    control = list(trace = 0)
  )

  if (!isTRUE(fit$convergence == 0)) {
    stop(sprintf("Rsolnp failed to converge, status %s.", fit$convergence))
  }

  w <- as.numeric(fit$pars)
  if (any(w < -1e-8) || any(w > 1 + 1e-8) || abs(sum(w) - 1) > 1e-7) {
    stop("Synthetic-control weights violate the simplex constraints.")
  }

  # Remove only machine-precision boundary noise, then restore exact adding up.
  w[abs(w) < 1e-12] <- 0
  w / sum(w)
}
```

We loop over the EU markets. For each one we take its three donors,
restrict to the years where the market and all three donors are observed,
solve for the weights, build the fitted path, and record the root mean
squared prediction error.

$$
\widehat{Y}_{j,t} = \sum_{k \in \Gamma_j} w_{jk}\,\log(\text{App}_{k,t}),
\qquad
\text{RMSPE}_j = \sqrt{\frac{1}{T_0}\sum_{t \le 2020}\bigl(\log(\text{App}_{j,t}) - \widehat{Y}_{j,t}\bigr)^2}
$$

```{r fit-loop}
#First block. Select the nearest donors only.

donor_selection <- data.frame()

for (j in eu_present) {
  
  d_row <- dist_treated_donors[j, ]
  d_sorted <- sort(d_row)
  donor_names <- names(d_sorted)[1:n_donors]
  
  for (k in 1:n_donors) {
    
    new_row <- data.frame(
      EU_country = j,
      donor_rank = k,
      donor = donor_names[k],
      distance = d_sorted[k],
      stringsAsFactors = FALSE
    )
    
    donor_selection <- rbind(donor_selection, new_row)
  }
}

rownames(donor_selection) <- NULL

head(donor_selection, n_donors)
```

```{r}
#Second block. Find usable years for each EU country and its donors.

year_selection <- data.frame()

for (j in eu_present) {
  
  donor_names <- donor_selection$donor[donor_selection$EU_country == j]
  selected_countries <- c(j, donor_names)
  
  block <- logmat[selected_countries, ]
  
  for (yr in colnames(block)) {
    
    values_this_year <- block[, yr]
    missing_values <- is.na(values_this_year)
    number_missing <- sum(missing_values)
    
    if (number_missing == 0) {
      
      new_row <- data.frame(
        EU_country = j,
        year = yr,
        usable = 1,
        stringsAsFactors = FALSE
      )
      
      year_selection <- rbind(year_selection, new_row)
    }
  }
}

rownames(year_selection) <- NULL

head(year_selection)
```

```{r}
#Third block. Estimate synthetic-control weights.

weight_results <- data.frame()

for (j in eu_present) {
  
  donor_names <- donor_selection$donor[donor_selection$EU_country == j]
  good_years <- year_selection$year[year_selection$EU_country == j]
  stopifnot(identical(sort(as.integer(good_years)), expected_years))
  
  y <- logmat[j, good_years]
  y <- as.numeric(y)
  
  Z <- logmat[donor_names, good_years]
  Z <- t(Z)
  
  w <- solve_weights(y, Z)
  
  for (k in 1:n_donors) {
    
    new_row <- data.frame(
      EU_country = j,
      donor = donor_names[k],
      synth_weight = w[k],
      stringsAsFactors = FALSE
    )
    
    weight_results <- rbind(weight_results, new_row)
  }
}

rownames(weight_results) <- NULL

head(weight_results, n_donors)
```

```{r}
fit_results <- data.frame()

for (j in eu_present) {
  
  donor_names <- donor_selection$donor[donor_selection$EU_country == j]
  good_years <- year_selection$year[year_selection$EU_country == j]
  
  y <- logmat[j, good_years]
  y <- as.numeric(y)
  
  Z <- logmat[donor_names, good_years]
  Z <- t(Z)
  
  w <- weight_results$synth_weight[weight_results$EU_country == j]
  
  fit <- Z %*% w
  fit <- as.numeric(fit)
  
  errors <- y - fit
  squared_errors <- errors^2
  mean_squared_error <- mean(squared_errors)
  rmspe <- sqrt(mean_squared_error)
  
  for (i in seq_along(good_years)) {
    
    new_row <- data.frame(
      EU_country = j,
      year = good_years[i],
      observed = y[i],
      synthetic = fit[i],
      error = errors[i],
      squared_error = squared_errors[i],
      rmspe = rmspe,
      stringsAsFactors = FALSE
    )
    
    fit_results <- rbind(fit_results, new_row)
  }
}

rownames(fit_results) <- NULL

head(fit_results)
```

```{r}
#Final block. Merge the donor information, distances, weights, and RMSPE.

donor_pool <- merge(
  donor_selection,
  weight_results,
  by = c("EU_country", "donor"),
  all.x = TRUE
)

rmspe_results <- unique(fit_results[, c("EU_country", "rmspe")])

donor_pool <- merge(
  donor_pool,
  rmspe_results,
  by = "EU_country",
  all.x = TRUE
)

donor_pool <- donor_pool[order(donor_pool$EU_country, donor_pool$donor_rank), ]

rownames(donor_pool) <- NULL

head(donor_pool, n_donors)
```

```{r}
# we create this for the batch 3 checks in the folder (we replicate the logic with EXCEL SOLVER)

#logmat_df <- data.frame(
#  country = rownames(logmat),
#  as.data.frame(logmat),
#  row.names = NULL,
#  check.names = FALSE
#)

#write_xlsx(
#  list(
#    logmat = logmat_df,
#    donor_selection = donor_selection,
#    weight_results = weight_results
#  ),
#  path = "solver_check_database.xlsx"
#)
```

# Aggregating to one global weight per donor

The difference-in-differences model uses a single common control path
for every EU market, so the country-specific weights have to be
collapsed into one weight per donor. We average each donor's weight
across the treated markets, treating a donor that never enters a pool as
a zero for that market.

$$
w_k = \frac{1}{J}\sum_{j=1}^{J} w_{jk}
$$

We then normalise so the global weights sum to one. Because each EU
market already contributed weights that summed to one, this
normalisation is a numerical safeguard and is exactly neutral up to
floating-point tolerance.

$$
\tilde{w}_k = \frac{w_k}{\sum_s w_s}
$$

Finally, to give the control side the same total influence as the
treated side in the weighted regression, the normalised donor weights
are scaled by the number of treated markets $J$, while every EU
observation carries unit weight.

$$
\omega_i =
\begin{cases}
1 & i \in \Omega \ \text{(treated EU)} \\
J \cdot \tilde{w}_k & i \in P \ \text{(donor)}
\end{cases}
$$

```{r global-weights}
J <- length(eu_present)   # number of treated markets that got a synthetic control
stopifnot(J == length(eu_real))

weights <- donor_pool %>%
  group_by(donor) %>%
  summarise(w_k = sum(synth_weight) / J, .groups = "drop") %>%
  mutate(
    normalized_weight  = w_k / sum(w_k),     # w-tilde
    regression_weight  = J * normalized_weight
  ) %>%
  rename(donor_country = donor, global_weight = w_k) %>%
  arrange(desc(normalized_weight))

weights

stopifnot(
  abs(sum(weights$normalized_weight) - 1) < 1e-10,
  abs(sum(weights$regression_weight) - J) < 1e-8
)
```

```{r write-excel}
write_xlsx(
  list(
    donor_pool = donor_pool,
    weights    = weights
  ),
  path = "control_group_output.xlsx"
)
```

## Reproducibility record

```{r session-info}
sessionInfo()
```
